---
title: "Building a curated Alaska Fisheries Science Center genetic sample repository to facilitate long-term use of molecular information for fisheries management"
author: ""
date: "11/29/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Summary
The Alaska Fisheries Science Center Genetics Program (AFSC GP) conducts 
research on Pacific salmon, groundfish, and forage fish species to comply 
with regional, national, and international agreements and treaties for 
managing marine resources. Much of this work relies on the collection of
samples, both fishery dependent and independent, to evaluate the existence 
and extent of population genetic structure in a variety of species. 
Currently, the genetics program manages a database of these samples, associated 
collection information, and laboratory data in FileMaker software. Filemaker is 
a relational database application with a graphical user interface. Recent 
limitations with the current database highlight the need for better records 
management. This project's primary goal is to work with the Alaska Fisheries 
Information Network (AKFIN) to develop a database model and
repository for AFSC GP's inventory of genetic samples. Additionally, a data 
management and upload tool in Oracle APEX will be created to interface 
with these data. These tables will be used for storing sample information 
including species, tissue type, collection information, and laboratory processing
data. Effective organization of these samples will aid in both the evaluation 
of population genetic trends across a diverse array of taxa as well as 
the development of molecular resources for fisheries
management. This project would bring consistency to AFSC GP's records 
management, as current salmon bycatch analyses rely heavily on AKFIN's database.

## Description of the current database
```{r RdInDat, eval=F, echo = T}
FMdat <- readxl::read_xlsx("./ABL_dat/20260127_export.xlsx")

FMupload <- readxl::read_xlsx("./ABL_dat/upload_template_fieldValidation.xlsx",
                              sheet='Field_Descriptions',
                              range='A1:B67',
                              col_names=F)
```

```{r FieldDescip, eval=T, echo=F}
FMupload %>%
  rename('Field' = 1,
         'Description' = 2) %>%
  knitr::kable()%>%
  kableExtra::kable_styling(full_width = F)

```

The current Filemaker database export (20260127) has a total of 68 fields.

```{r Salmon, eval=F, echo = T}
Salmon_Entries <- FMdat %>%
  filter(grepl(pattern='Salmonidae',FamilyOLD) | grepl(pattern='King|king|Chinook|chinook|Pink|pink|Coho|coho|Chum|chum|salmon|Steelhead|steelhead',CommonNameOLD)|grepl(pattern='Oncorhynchus',GenusOLD)|grepl(pattern='Oncorhynchus',SpeciesNameOLD))
  
```

There are 54,596 entries, of which 18,355 are salmon (33.6%). I am going to focus on 
cleaning up the Salmon entries. 

## Salmoniformes Entries

```{r Genera, eval=T, echo=T}
#Break into ind. species
Salmon_Entries %>% 
  select(GenusOLD) %>%
  table()

```

```{r NonOncorhynchus, eval=T, echo=T}
Salmon_Entries %>% 
  filter(GenusOLD != 'Oncorhynchus') %>%
  select(SpeciesNameOLD)%>%
  table()

```

Paralithodes camtschaticus is king crab, Lamna ditropis is 
a salmon shark, and Coregonus clupeaformis is lake whitefish which are Salmonidae... so
we'll keep those. 

```{r rm2Sp, eval=T, echo=F}
Salmon_Entries <- Salmon_Entries %>%
  filter(!SpeciesNameOLD %in% c('Lamna ditropis', 'Paralithodes camtschaticus'))
```

#### Common Names
Now lets look at the common names. We can standardize these and then
confirm that the genus and species names are correct. 

```{r CommonNames, eval=T, echo=T}
Salmon_Entries %>% 
  select(CommonNameOLD)%>%
  table()
```

That's unfortunate, looks like there are 5 entries for chinook salmon alone. 

```{r StndrzCN, eval=T, echo=T}
Salmon_Entries <- Salmon_Entries %>%
  mutate(CommonNameOLDlc = tolower(CommonNameOLD))%>%
  mutate(CommonName = recode(CommonNameOLDlc, 
                'chinook' = 'Chinook salmon',
                'king salmon' = 'Chinook salmon',
                'chinook fin clip' = 'Chinook salmon',
                'chinook (bycatch)' = 'Chinook salmon',
                'chinook salmon' = 'Chinook salmon',
                'chinook/king salmon' = 'Chinook salmon'))

Salmon_Entries %>% 
  select(CommonName)%>%
  table()

Salmon_Entries %>% nrow() == (Salmon_Entries %>% 
                                select(CommonName)%>%
                                  table() %>% sum())
```

Dang, some of the common names are abscent. Hopefully there is an entry in 
the genus and species. 

```{r NoCN, eval=T, echo=T}
NoCommonNames <- Salmon_Entries %>%
                   filter(is.na(CommonName))

Salmon_Entries <- Salmon_Entries %>%
  mutate(CommonName = ifelse(grepl(pattern='CMGIS04|CMPEL93',AlternateID)&is.na(CommonName),'chum salmon', CommonName))

```

Nope, but at least they were ADFG collections. In the Silly code (Alt ID), 
chum collections start with CM, king salmon with K, sockeye with S. So
at least we know that these are chum salmon. 

### Genus, Family, and Species
Now that all the common names are filled in we can work with the genus name.

```{r StndrzGenus, eval=T, echo=T}
Salmon_Entries %>%
  select(GenusOLD) %>%
  table
  
#Damn very few of them are filled out.  

Salmon_Entries <- Salmon_Entries %>%
   mutate(Genus = ifelse(grepl(pattern='Chinook|pink|coho|chum|sockeye',CommonName),'Oncorhynchus',
                               ifelse(grepl(pattern='lake whitefish',CommonName),'Coregonus',NA)))

Salmon_Entries %>%
  select(Genus) %>%
  table
```

```{r StndrzFamily, eval=T, echo=T}
Salmon_Entries %>%
  select(FamilyOLD) %>%
  table
  
Salmon_Entries %>% nrow() == (Salmon_Entries %>% 
                                select(FamilyOLD)%>%
                                  table() %>% sum())

```

```{r StndrzSpecies, eval=T, echo=T}
Salmon_Entries %>%
  select(SpeciesNameOLD) %>%
  table
  
#It appears that in the Field_Values validation we don't
#want the binomial latin name as species, just the species 

Salmon_Entries <- Salmon_Entries %>%
  mutate(Species = recode(SpeciesNameOLD, 
                'Coregonus clupeaformis' = 'clupeaformis',
                'Onchorhynchus gorbuscha' = 'gorbuscha',
                'Onchorynchus tshawytscha' = 'tshawytscha',# Oncorhynchus misspelled
                'Oncorhynchus kisutch' = 'kisutch',
                'Oncorhynchus nerka' = 'nerka',
                'Oncorhynchus tshawytscha' = 'tshawytscha'))

Salmon_Entries %>%
  select(Species) %>%
  table

NoSpeciesName <- Salmon_Entries %>%
  filter(is.na(Species))

NoSpeciesName %>%
  select(CommonName) %>%
  table()

Salmon_Entries <- Salmon_Entries %>%
  mutate(Species = ifelse((is.na(Species)&CommonName == "Chinook salmon"),'tshawytscha',Species))%>%
  mutate(Species = ifelse(is.na(Species)&CommonName == "chum salmon",'keta',Species))%>%
  mutate(Species = ifelse(is.na(Species)&CommonName == "coho salmon",'kisutch',Species))%>%
  mutate(Species = ifelse(is.na(Species)&CommonName == "pink salmon",'gorbuscha',Species))%>%
  mutate(Species = ifelse(is.na(Species)&CommonName == "sockeye salmon",'nerka',Species))%>%
  mutate(Species = ifelse(is.na(Species)&CommonName == "lake whitefish",'clupeaformis',Species))
  
Salmon_Entries %>% nrow() == (Salmon_Entries %>% 
                                select(Species)%>%
                                  table() %>% sum())

Salmon_Entries %>%
  select(Species) %>%
  table

```

Sweet, those look pretty good for species names. 

### Contact, Donating Lab, Collector

In the field descriptions 'Contact' is listed as a required field with 
instructions about field validations; however there are values listed in
the sheet. Additionally, contact is not listed in the 'ABLG-DB-working-copy'
sheet for upload. So perhaps that field has been replaced with 
'Collector' which has 69 unique values. 

```{r StndrzCollector, eval=T, echo=T}
Salmon_Entries %>%
  select(Collector) %>%
  table
  
Salmon_Entries <- Salmon_Entries %>%
  mutate(Collector = recode(Collector,
                            'LPW-Waters' = 'Charlie Waters',
                            'Vulstek' = 'Scott Vulstek',
                            'Mike-Courtney' = 'Michael Courtney'))%>%
  mutate(Collector = ifelse(Collector=="---",NA,Collector))

NoCollector <- Salmon_Entries %>%
  filter(is.na(Collector)) 

```

There are 10,941 records without a collector recorded. Maybe that is fine provided
we have a contact/donating lab for the sample. 

Now take a stab at the Donating Lab

```{r StndrzCollector, eval=T, echo=T}
Salmon_Entries %>%
  select(DonatingLab) %>%
  table
  
Salmon_Entries <- Salmon_Entries %>%
  mutate(DonatingLab = recode(DonatingLab,
                            'LPW/Waters' = 'AFSC-LPW',
                            'Little Port Walter' = 'AFSC-LPW',
                            'Vulstek' = 'AFSC-Scott Vulstek',
                            'Vulstek?' = 'AFSC-Scott Vulstek',
                            'Seitz' = 'UAF-Andy Seitz',
                            'Manhard Mol Bio 2018' = 'UAF-Anthony Gharrett'))%>%
  mutate(DonatingLab = ifelse(DonatingLab=="---",NA,DonatingLab))

Salmon_Entries %>%
  select(DonatingLab) %>%
  table

sum(is.na(Salmon_Entries$DonatingLab))

```

```{r UnknwnRcd, eval=T, echo=T}
BadRcrds <- Salmon_Entries %>%
  filter(is.na(DonatingLab)&is.na(Collector))
```

What are the samples from 

BC22_25300_5006_des/env Chinook bycatch?

```{r}
# MCGL_18_ - Coregonus clupeaformis? whitefish_01 plate; Jared Homola Molecular Conservation Genetics Lab
Salmon_Entries <- Salmon_Entries %>%
  mutate(DonatingLab = ifelse(is.na(DonatingLab) & CommonName == 'lake whitefish','MCGL-Jared Homola',DonatingLab))

#Sckye_2014_Tog_XX - Little Togiak Beach; Daniel Schindler UW
Salmon_Entries <- Salmon_Entries %>%
  mutate(DonatingLab = case_when(
    is.na(DonatingLab) & grepl("Sckye_2014_Tog_", AlternateID) ~ "UW-Daniel Schindler",
    TRUE ~ DonatingLab # 
  ))

#AFSC-Ingrid Spies 
# PL_XX - Sckye_LkWA2001
# 11/7/00_CedarSockeye_XX - Sckye_Cedar_1999_2000
# Cedar_R_1999_ - 
# Sckye2001_Baker_

Salmon_Entries <- Salmon_Entries %>%
  mutate(DonatingLab = case_when(
    is.na(DonatingLab) & grepl("11/7/00_CedarSockeye_|Sckye_Cedar_1999_2000|Cedar_R_1999_|Sckye_Baker_|Sckye2001_Baker_", AlternateID) ~ "AFSC-Ingrid Spies",
    TRUE ~ DonatingLab # 
  ))

Salmon_Entries <- Salmon_Entries %>%
  mutate(DonatingLab = case_when(
    is.na(DonatingLab) & grepl("Sckye_LkWA2001", ExtractionPlate) ~ "AFSC-Ingrid Spies",
    TRUE ~ DonatingLab # 
  ))


# DipacMale/FemaleCoho_2011; DIPAC-Katie Harms
Salmon_Entries <- Salmon_Entries %>%
  mutate(DonatingLab = case_when(
    is.na(DonatingLab) & grepl("Coho_2011", AlternateID) ~ "DIPAC-Katie Harms",
    TRUE ~ DonatingLab # 
  ))


# BC22_25300_5006; NMFS-Observer Program
Salmon_Entries <- Salmon_Entries %>%
  mutate(DonatingLab = case_when(
    is.na(DonatingLab) & grepl("BC22_", AlternateID) ~ "NMFS-Observer Program",
    TRUE ~ DonatingLab # 
  ))
```



#### Collection Year, Month, Day

```{r Year, eval=T, echo=F}
Salmon_Entries %>%
  select(CollectionYear) %>%
  unlist()%>%
  table
  
NoYear <- Salmon_Entries %>%
  filter(is.na(CollectionYear))
```

```{r Month, eval=T, echo=F}
Salmon_Entries %>%
  select(CollectionMonth) %>%
  unlist()%>%
  table
 
Salmon_Entries %>%
  filter(CollectionMonth == 'NA')

Salmon_Entries <- Salmon_Entries %>%
  mutate(CollectionMonth = ifelse(CollectionMonth=="NA",NA,CollectionMonth))

NoMonth <- Salmon_Entries %>%
  filter(is.na(CollectionMonth))
```

```{r Year, eval=T, echo=F}
Salmon_Entries %>%
  select(CollectionDay) %>%
  unlist()%>%
  table
  
ShitSamp <- Salmon_Entries %>%
  filter(CollectionDay=='103') %>%
  select(ABLG,AlternateID,CommonName,Genus,Species,CollectionYear,CollectionMonth,CollectionDay,Location1,Location2,FreshwaterWatershed)

Salmon_Entries %>%
  filter(ABLG %in% c(41646,41647,41648,41649,41650)) %>%
  select(ABLG,AlternateID,CommonName,Genus,Species,CollectionYear,CollectionMonth,CollectionDay,Location1,Location2,FreshwaterWatershed)

#Turns out that Little Whitefish should be 8/11
Salmon_Entries %>%
  filter(grepl(pattern='SLWHI11',AlternateID))%>%
  select(ABLG,AlternateID,CommonName,Genus,Species,CollectionYear,CollectionMonth,CollectionDay,Location1,Location2,FreshwaterWatershed)

Salmon_Entries <- Salmon_Entries %>%
  mutate(CollectionDay = ifelse(grepl(pattern='SLWHI11',AlternateID),'11',CollectionDay))
  
Salmon_Entries <- Salmon_Entries %>%
  mutate(CollectionDay = recode(CollectionDay,
                                '1' = '01',
                                '2' = '02',
                                '3' = '03',
                                '4' = '04',
                                '5' = '05',
                                '6' = '06',
                                '7' = '07',
                                '8' = '08',
                                '9' = '09'))

Salmon_Entries %>%
  select(CollectionDay) %>%
  unlist()%>%
  table
```


### Location

```{r MR, eval=T, echo=F}
Salmon_Entries %>%
  select(MarineRegion) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(MarineRegion = ifelse(MarineRegion=="---",NA,MarineRegion))

```

```{r FWWS, eval=T, echo=F}
Salmon_Entries %>%
  select(FreshwaterWatershed) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(FreshwaterWatershed = ifelse(FreshwaterWatershed=="---",NA,FreshwaterWatershed))

```

```{r Loc1, eval=T, echo=F}
Salmon_Entries %>%
  select(Location1) %>%
  unlist()%>%
  table

Chickamin <- Salmon_Entries %>%
  filter(grepl(pattern='CHICKAMIN',Location1))%>%
  select(ABLG,AlternateID,CommonName,Genus,Species,CollectionYear,CollectionMonth,CollectionDay,Location1,Location2,FreshwaterWatershed)

Salmon_Entries <- Salmon_Entries %>%
  mutate(Location1 = gsub(pattern='"Replace with',replacement="",Location1))%>%
  mutate(Location1 = gsub(pattern='CHICKAMIN',replacement="Chickamin",Location1))%>%
  mutate(Location1 = gsub(pattern='ANDREW',replacement="Andrew",Location1))%>%
  mutate(Location1 = ifelse(Location1=="---",NA,Location1))%>%
  mutate(Location1 = str_trim(Location1))

```

```{r Loc2, eval=T, echo=F}
Salmon_Entries %>%
  select(Location2) %>%
  unlist()%>%
  table

#Probably below the wier samples
AukeFalls <- Salmon_Entries %>%
  filter(grepl(pattern='Auke Cr. Falls',Location2))%>%
  select(ABLG,AlternateID,CommonName,Genus,Species,CollectionYear,CollectionMonth,CollectionDay,Location1,Location2,FreshwaterWatershed)

Salmon_Entries <- Salmon_Entries %>%
  mutate(Location2 = ifelse(Location2=="---",NA,Location2))%>%
  mutate(Location2 = str_trim(Location2))

```


```{r Lat, eval=T, echo=F}
Salmon_Entries %>%
  select(StartLatitudeDD) %>%
  unlist()%>%
  table

#0 isn't valid
Salmon_Entries <- Salmon_Entries %>%
  mutate(StartLatitudeDD = ifelse(StartLatitudeDD==0,NA,StartLatitudeDD))

Salmon_Entries %>%
  select(EndLatitudeDD) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(EndLatitudeDD = ifelse(EndLatitudeDD=="---",NA,EndLatitudeDD))
```

```{r Long, eval=T, echo=F}
Salmon_Entries %>%
  select(StartLongitudeDD) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(EndLongitudeDD) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(EndLongitudeDD = ifelse(EndLongitudeDD=="---",NA,EndLongitudeDD))
```

```{r Long, eval=T, echo=F}
Salmon_Entries %>%
  select(EstimatedLatLongYesNo) %>%
  unlist()%>%
  table
```

The estimated or not estimated lat/long for each collection should be 
entered with the metadata, looks like I might have to go back and 
fill these in for salmon collections. 

### Survey Specific Fields maybe not super relevant to salmon

```{r Survey, eval=T, echo=F}
Salmon_Entries %>%
  select(CruiseID) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(HaulID) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(StationID) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(Equipment_Gear) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(DepthM) %>%
  unlist()%>%
  table
#is 0 valid for depth in meters?

Salmon_Entries %>%
  select(StartDepthM) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(EndDepthM) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(TempC) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(Habitat) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(CruiseID = ifelse(CruiseID=="---",NA,CruiseID))%>%
  mutate(HaulID = ifelse(HaulID=="---",NA,HaulID))%>%
   mutate(StationID = ifelse(StationID=="---",NA,StationID))%>%
  mutate(Habitat = ifelse(Habitat=="---",NA,Habitat))%>%
   mutate(Equipment_Gear = ifelse(Equipment_Gear=="---",NA,Equipment_Gear))%>%
  mutate(DepthM = ifelse(DepthM==0,NA,DepthM))%>%
  mutate(TempC = ifelse(TempC==0,NA,TempC))

```

#### Morphometrics

```{r Morpho, eval=T, echo=F}
Salmon_Entries %>%
  select(LengthMM) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(LengthMM = ifelse(LengthMM==0,NA,LengthMM))

Salmon_Entries %>%
  select(LengthType) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(FishDepthMM) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(FishDepthMM = ifelse(FishDepthMM %in% c(T,F),NA,FishDepthMM))


Salmon_Entries %>%
  filter(!is.na(FishDepthMM)) %>%
select(ABLG,AlternateID,CommonName,Genus,Species,CollectionYear,CollectionMonth,CollectionDay,Location1,Location2,FreshwaterWatershed)


Salmon_Entries %>%
  filter(!is.na(FishDepthMM)) %>%
select(ABLG,AlternateID,CommonName,CollectionYear,CollectionMonth,Location1,FishDepthMM)


Salmon_Entries %>%
  select(WeightG) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(WeightG = ifelse(WeightG==0,NA,WeightG))

#That 17,940g fish is 39.5lbs!


Salmon_Entries %>%
  select(Lifestage) %>%
  unlist()%>%
  table

#These are all ages, what the hell is 1.10000000001, 2.20000002, 2.29999998. Need to go back to the actual data for those samples
# should this be more of a categorical lifestage not age? smolt, juvenile, adult etc? We have age in years
Salmon_Entries <- Salmon_Entries %>%
  mutate(Lifestage = ifelse(Lifestage=="---",NA,Lifestage))

StrngLS <- Salmon_Entries %>%
  filter(Lifestage %in% c("1.1000000000000001","2.2000000000000002","2.2999999999999998"))%>%
  select(ABLG,AlternateID,CommonName,CollectionYear,CollectionMonth,Location1,Lifestage,AgeYRS)

#After looking at the sockeye metadata it appears that there was an excel file mixup that lead to 
#issues with the formatting of the European ages for some salmon records
Salmon_Entries <- Salmon_Entries %>%
  mutate(Lifestage = case_when(
    Lifestage == "1.1000000000000001" ~ "1.1",
    Lifestage == "2.2000000000000002" ~ "2.2",
    Lifestage == "2.2999999999999998" ~ "2.3",
    TRUE ~ as.character(Lifestage) 
  ))

Salmon_Entries <- Salmon_Entries %>%
  mutate(AgeYRS = case_when(
    AgeYRS == "1.1000000000000001" ~ "1.1",
    TRUE ~ as.character(AgeYRS) 
  ))

# Now lets make a column for Age_Total and Age_European
EuropeanAges <- Salmon_Entries %>%
  filter(grepl(pattern="\\.",Lifestage)|grepl(pattern="\\.",AgeYRS))

Salmon_Entries <- Salmon_Entries %>%
  mutate(Age_European = ifelse(grepl(pattern="\\.",Lifestage),Lifestage,
                                 ifelse(grepl(pattern="\\.",AgeYRS), AgeYRS,
                                        NA)))

# And then convert all the ages in Lifestage to Adult, cuase they were sampled 
# at a weir and were adults of different ages.
Salmon_Entries <- Salmon_Entries %>%
  mutate(Lifestage = case_when(
    grepl("\\.", Lifestage) ~ 'adult',
    grepl("jack", Lifestage) ~ 'adult',
    TRUE                    ~ Lifestage))


Salmon_Entries %>%
  select(Lifestage) %>%
  unlist()%>%
  table

#good now I just have adult and juvenile...  
Juvies <- Salmon_Entries %>% filter(Lifestage == 'juvenile')

Salmon_Entries %>%
  select(AgeYRS) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(AgeYRS = ifelse(AgeYRS=="---",NA,AgeYRS))

Salmon_Entries <- Salmon_Entries %>%
  mutate(Age_European = na_if(Age_European, 'r.r'))%>%
  mutate(Age_European = na_if(Age_European, 'r.1'))
    

AmbigAges <- Salmon_Entries %>% filter(grepl(pattern="or",AgeYRS))

Salmon_Entries <- Salmon_Entries %>%
  mutate(AgeYRS = case_when(
    grepl("or", AgeYRS) ~ NA,
    TRUE ~ AgeYRS))

 Salmon_Entries <- Salmon_Entries %>%
  mutate(Age_Total = case_when(
    grepl("\\.", Age_European) ~ map_dbl(str_split(Age_European, "\\."), ~ sum(as.numeric(.x))+1),
    TRUE                      ~ as.numeric(AgeYRS)))


 Salmon_Entries %>%
  select(Age_Total) %>%
  unlist()%>%
  table
 
 Salmon_Entries %>%
  select(Age_European) %>%
  unlist()%>%
  table
 
 
Salmon_Entries %>%
  select(Sex) %>%
  unlist()%>%
  table

Salmon_Entries %>% filter(grepl(pattern='1|2|3',Sex)) %>% select(ABLG,AlternateID,DonatingLab,CommonNameOLD)
#1,2,3 samples are from ADFG so 1 = Male, 2 = Female, and 3 = NA


Salmon_Entries <- Salmon_Entries %>%
  mutate(Sex = case_when(
    Sex=="---" ~ NA,
    Sex == 'F' ~ 'female',
    Sex == 'M' ~ 'male',
    Sex == 'male (jack)' ~ 'jack',
    Sex == 'U' ~ NA,
    Sex == '3' ~ NA,
    Sex == '1' ~ 'male',
    Sex == '2' ~ 'female'))
  
Salmon_Entries %>%
  select(CaptureTime) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(ReleaseTime) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(IdentifiedBy) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(IdentifiedBy = ifelse(IdentifiedBy=="---",NA,IdentifiedBy))

```

#### Lab fields

```{r Survey, eval=T, echo=F}
Salmon_Entries %>%
  select(Preservation) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(Preservation = ifelse(Preservation=="---",NA,Preservation))%>%
  mutate(Preservation = recode(Preservation,
                               'dried tissue' = 'Dried',
                               'whatman' = 'Dried',
                               'Dry' = 'Dried',
                               'ethanol' = 'Ethanol')) #based on Validation keys in template

Salmon_Entries %>%
  select(SampleRoom) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(SampleRoom = recode(SampleRoom,
                              'Gave back to Wes (to give back to Charlie?)' = 'returned to C. Waters',
                               'icnthyology' = 'Ichthyology (L107)',
                               'Ichthy' = 'Ichthyology',
                               'Ichthyology' = 'Ichthyology (L107)'))

Salmon_Entries <- Salmon_Entries %>%
  mutate(SampleRoom = ifelse(SampleRoom=="---",NA,SampleRoom))%>%
  mutate(SampleRoom = ifelse(SampleRoom=="NA",NA,SampleRoom))
  

Salmon_Entries %>%
  select(SampleLocation) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(SampleLocation = ifelse(SampleLocation=="---",NA,SampleLocation))%>%
  mutate(SampleLocation = ifelse(SampleLocation=="NA",NA,SampleLocation))

Salmon_Entries %>%
  select(SampleSublocation) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(SampleSublocation = ifelse(SampleSublocation=="---",NA,SampleSublocation))

#not sure if any of these are valid sublocations? LPW ask Juliana

Salmon_Entries %>%
  select(ExtractedBy) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(ExtractedBy = ifelse(ExtractedBy=="Juliana","Juliana Cornett",ExtractedBy))%>%
  mutate(ExtractedBy = ifelse(ExtractedBy=="---",NA,ExtractedBy))


Salmon_Entries %>%
  select(ExtractionDate) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(ExtractionDate = ifelse(ExtractionDate=="---",NA,ExtractionDate))%>%
  mutate(ExtractionDate = recode(ExtractionDate, #format YYYYMMDD
                                 '10/25/2022' = '20221025',
                                 '12/19/2022' = '20221219', 
                                 '3/30/2021' = '20210330',
                                 '5/16/2021' = '20210516', 
                                 '5/23/2021' = '20210523',
                                 '5/24/2022' = '20220524',
                                 '5/4/2021' = '20210504',
                                 '8/18/2022' = '20220818',
                                 '9/21/2022' = '20220921'))

Salmon_Entries %>%
  select(ExtractionMethod) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(ExtractionMethod = recode(ExtractionMethod, 
                                 'QIAGEN DNeasy' = 'QIAGEN',
                                 'Macherey Nagel' = 'Macherey-Nagel',
                                 'Machery-Nagel' = 'Macherey-Nagel'))%>%
  mutate(ExtractionMethod= ifelse(ExtractionMethod=="Unknown",NA,ExtractionMethod))


Salmon_Entries %>%
  select(ExtractionPlate) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(ExtractionPlate = ifelse(ExtractionPlate=="---",NA,ExtractionPlate))

Salmon_Entries %>%
  select(ExtractionWell) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(ExtractionWell = ifelse(ExtractionWell=="---",NA,ExtractionWell))

#Need to decide if we want to have A01 or A1 for well locations and be consistent

Salmon_Entries %>%
  select(DNAngul) %>%
  unlist()%>%
  table

#0s could be valid here.

Salmon_Entries %>%
  select(DNAquality) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(DNAquality = ifelse(DNAquality=="---",NA,DNAquality))%>%
  mutate(DNAquality = str_trim(DNAquality))%>%
  mutate(DNAquality = recode(DNAquality,
                             'HMW band and smearing' = 'HMW bands and smearing',
                             'No HMW bands and no smearing' = 'no HMW bands and no smearing',
                             'no HMW bands and no smearing (empty)' = 'no HMW bands and no smearing',
                             'no HMW bands and no smearing (empty) (empty)' = 'no HMW bands and no smearing',
                             'No HMW bands and smearing' = 'no HMW bands and smearing'))


Salmon_Entries %>%
  select(DNAroom) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(DNAroom = ifelse(DNAroom=="---",NA,DNAroom))

Salmon_Entries %>%
  select(DNAlocation) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(DNAlocation = ifelse(DNAlocation=="---",NA,DNAlocation))%>%
  mutate(DNAlocation = recode(DNAlocation,
                              'F212_1' = 'F212-1',
                              'F213_7' = 'F213-7'))


Salmon_Entries %>%
  select(DNAsublocation) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(DNAsublocation = ifelse(DNAsublocation=="---",NA,DNAsublocation))

#WGS1
Salmon_Entries %>%
  select(WGSDateYYYYMMDD) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(WGSBatchID) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(WGSLaneID) %>%
  unlist()%>%
  table

#WGS2 -None have been used multiple times, all blanks
Salmon_Entries %>%
  select(WGSDate2YYYYMMDD) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(WGSBatchID2) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(WGSLaneID2) %>%
  unlist()%>%
  table

#GTseq - All left empty... just not updated. 
Salmon_Entries %>%
  select(GTSeqDateYYYYMMDD) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(GTSeqPanel) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(GTSeqDate2YYYYMMDD) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(GTSeqPanel2) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(GTSeqDate3YYYYMMDD) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(GTSeqPanel3) %>%
  unlist()%>%
  table

#Genetic ID missing

#Comments are fine?

Salmon_Entries <- Salmon_Entries %>%
  select(-c(CommonNameOLDlc,CommonNameOLD,GenusOLD,SpeciesNameOLD))%>%
  relocate(CommonName,.after='AlternateID')%>%
  relocate(Genus,.after='FamilyOLD')%>%
  relocate(Species,.after='Genus')%>%
  rename('Family' = 'FamilyOLD')


```

```{r SplitDonatingLab}
Salmon_Entries$DonatingLab %>% unique()

Salmon_Entries <- Salmon_Entries %>%
  mutate(DonatingOrganization = case_when(
   DonatingLab=="UW-Daniel Schindler" ~ "UW",
   DonatingLab=="Lorenz Hauser" ~ "UW",
   DonatingLab %in% c("AFSC-Ingrid Spies","AFSC-Scott Vulstek","AFSC-LPW") ~ "AFSC",
   DonatingLab %in% c("ADFG - Heather Hoyt","ADFG","ADFG-Anchorage") ~ "ADFG",
   DonatingLab=="SWFSC" ~ "SWFSC",
   DonatingLab=="ABL" ~ "AFSC",
   DonatingLab=="DIPAC-Katie Harms" ~ "DIPAC",
   DonatingLab=="NMFS-Observer Program" ~ "NMFS Observer Program",
   DonatingLab=="MCGL-Jared Homola" ~ "MCGL",
   DonatingLab %in% c("UAF-Anthony Gharrett","UAF" ,"UAF-Andy Seitz" ) ~ "UAF",
   DonatingLab=="MN-DNR-CGoldsworthy" ~ "MN DNR",
   DonatingLab=="Cory Dick" ~ "Humboldt State University",#Cory Dick
   DonatingLab=="Wes Larson" ~ "AFSC", #Wes Larson
   TRUE ~ DonatingLab))%>%
   mutate(DonatingContactName = case_when(
   DonatingLab=="UW-Daniel Schindler" ~ "Daniel Schindler",
   DonatingLab=="Lorenz Hauser" ~ "Lorenz Hauser",
   DonatingLab == "AFSC-Ingrid Spies" ~ "Ingrid Spies",
   DonatingLab == "AFSC-Scott Vulstek" ~ "Scott Vulstek",
   DonatingLab == "ADFG - Heather Hoyt" ~ "Heather Hoyt",
   DonatingLab=="DIPAC-Katie Harms" ~ "Katie Harms",
   DonatingLab=="MCGL-Jared Homola" ~ "Jared Homola",
   DonatingLab == "UAF-Anthony Gharrett" ~ "Anthony Gharrett",
   DonatingLab == "UAF-Andy Seitz" ~ "Andy Seitz",
   DonatingLab=="MN-DNR-CGoldsworthy" ~ "Cory Goldsworthy",
   DonatingLab=="Cory Dick" ~ "Cory Dick",#Cory Dick
   DonatingLab=="Wes Larson" ~ "Wes Larson", #Wes Larson
   TRUE ~ NA))

```


```{r WriteOutFile}
write_delim(x=Salmon_Entries %>%
              select(-CollectionDate)%>%#remove collection Date
              select(-DonatingLab), #remove DonatingLab
          file="Salmon_dbEntries_Clean.txt",
          delim="\t")

```


