---
title: "Building a curated Alaska Fisheries Science Center genetic sample repository to facilitate long-term use of molecular information for fisheries management"
author: ""
date: "11/29/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Summary
The Alaska Fisheries Science Center Genetics Program (AFSC GP) conducts 
research on Pacific salmon, groundfish, and forage fish species to comply 
with regional, national, and international agreements and treaties for 
managing marine resources. Much of this work relies on the collection of
samples, both fishery dependent and independent, to evaluate the existence 
and extent of population genetic structure in a variety of species. 
Currently, the genetics program manages a database of these samples, associated 
collection information, and laboratory data in FileMaker software. Filemaker is 
a relational database application with a graphical user interface. Recent 
limitations with the current database highlight the need for better records 
management. This project's primary goal is to work with the Alaska Fisheries 
Information Network (AKFIN) to develop a database model and
repository for AFSC GP's inventory of genetic samples. Additionally, a data 
management and upload tool in Oracle APEX will be created to interface 
with these data. These tables will be used for storing sample information 
including species, tissue type, collection information, and laboratory processing
data. Effective organization of these samples will aid in both the evaluation 
of population genetic trends across a diverse array of taxa as well as 
the development of molecular resources for fisheries
management. This project would bring consistency to AFSC GP's records 
management, as current salmon bycatch analyses rely heavily on AKFIN's database.

## Description of the current database
```{r RdInDat, eval=F, echo = T}
FMdat <- readxl::read_xlsx("./ABL_dat/20260127_export.xlsx")

FMupload <- readxl::read_xlsx("./ABL_dat/upload_template_fieldValidation.xlsx",
                              sheet='Field_Descriptions',
                              range='A1:B67',
                              col_names=F)
```

```{r FieldDescip, eval=T, echo=F}
FMupload %>%
  rename('Field' = 1,
         'Description' = 2) %>%
  knitr::kable()%>%
  kableExtra::kable_styling(full_width = F)

```

The current Filemaker database export (20260127) has a total of 68 fields.

Pat will take care of the salmon and Diana will address the rockfish.

```{r Salmon, eval=F, echo = T}
Salmon_Entries <- FMdat %>%
  filter(grepl(pattern='Salmonidae',FamilyOLD) | grepl(pattern='King|king|Chinook|chinook|Pink|pink|Coho|coho|Chum|chum|salmon|Steelhead|steelhead',CommonNameOLD)|grepl(pattern='Oncorhynchus',GenusOLD)|grepl(pattern='Oncorhynchus',SpeciesNameOLD)) %>%
  filter(!SpeciesNameOLD %in% c('Lamna ditropis', 'Paralithodes camtschaticus'))

Rockfish_Entries <- FMdat %>%
  filter(
    grepl(pattern='Sebastidae',FamilyOLD) |
      grepl(pattern='Rockfish|rockfish',CommonNameOLD)|
      grepl(pattern='Sebastes|SEBASTES|Sebastolobus|Sebastulobus',GenusOLD)|
      grepl(pattern='Sebastes|Sebastolobus',SpeciesNameOLD))

Other_Entries <- FMdat %>%
  anti_join(Salmon_Entries) %>%
  anti_join(Rockfish_Entries)
```


## Not Salmoniformes or Sebastes/Sebastolobus Entries

check what's there 
```{r Genera, eval=T, echo=T}
Other_Entries %>% 
  select(GenusOLD) %>%
  table()
```

```{r}
Other_Entries %>% 
  select(CommonNameOLD)%>%
  table()
```

```{r}
Other_Entries %>% 
  select(SpeciesNameOLD)%>%
  table()
```


#### Common Names

I'm going to follow Pat and first standardize the common names.
```{r StndrzCN, eval=T, echo=T}
Other_Entries <- Other_Entries %>%
  mutate(CommonNameOLD = as.character(CommonNameOLD),
         CommonNameOLD = stringr::str_trim(CommonNameOLD),
         CommonNameOLD = stringr::str_to_lower(CommonNameOLD),
         CommonName = case_when(
           stringr::str_detect(CommonNameOLD, "pollock") ~ "walleye pollock",
           stringr::str_detect(CommonNameOLD, "pacific cod") ~ "pacific cod",
           stringr::str_detect(CommonNameOLD, "channel catfish") ~ "channel catfish",
           stringr::str_detect(CommonNameOLD, "largemouth bass") ~ "largemouth bass",
           stringr::str_detect(CommonNameOLD, "pacific sardine") ~ "pacific sardine",
           stringr::str_detect(CommonNameOLD, "steller sea lion") ~ "steller sea lion",
      # Default: keep original
      TRUE ~ CommonNameOLD
    )
  )

#what are 'crab', 'placeholder-unknown', 'seal scat', 'sleeper shark ?'
Other_Entries %>% filter(CommonNameOLD == 'sleeper shark ?')

Other_Entries %>% 
  select(CommonName)%>%
  table()

Other_Entries %>% nrow() == (Other_Entries %>% 
                                select(CommonName)%>%
                                  table() %>% sum())
```

Dang, some of the common names are absent. Hopefully there is an entry in 
the genus and species. 

```{r NoCN, eval=T, echo=T}
NoCommonNames <- Other_Entries %>%
                   filter(is.na(CommonName))

NoCommonNames %>% 
  select(SpeciesNameOLD)%>%
  table()
```

add common name for those that we can 
```{r}
Other_Entries <- Other_Entries %>%
  mutate(
    CommonName = case_when(
      SpeciesNameOLD == 'Arctogadus glacialis' ~ 'polar cod',
      SpeciesNameOLD == 'Boreogadus saida' ~ 'arctic cod',
      SpeciesNameOLD == 'Clupea pallasii' ~ 'pacific herring',
      SpeciesNameOLD == 'Eleginus gracilis' ~ 'saffron cod',
      SpeciesNameOLD == 'Gadus chalchogrammus' ~ 'walleye pollock',
      SpeciesNameOLD == 'Gadus chalcogrammus' ~ 'walleye pollock',
      SpeciesNameOLD == 'Gadus macrocephalus' ~ 'pacific cod',
      SpeciesNameOLD == 'Hippoglossoides elassodon' ~ 'flathead sole',
      SpeciesNameOLD == 'Hippoglossoides robustus' ~ 'bering flounder',
      SpeciesNameOLD == 'Limanda aspera' ~ 'yellowfin sole',
      SpeciesNameOLD == 'Microgadus proximus' ~ 'pacific tomcod',
      TRUE ~ CommonName
    )
  )
```

```{r}
NoCommonNames <- Other_Entries %>%
                   filter(is.na(CommonName))

NoCommonNames %>% 
  select(SpeciesNameOLD)%>%
  table()
```

also some missing ID's for juvenile cod from Johanna here.


### Genus, Family, and Species
Now that *most* the common names are filled in we can work with the genus name.

```{r StndrzGenus, eval=T, echo=T}
Other_Entries %>%
  select(GenusOLD) %>%
  table
#incomplete and incorrect in some cases (Pollachius)  

Other_Entries <- Other_Entries %>%
  mutate(GenusOLD = as.character(GenusOLD),
         GenusOLD = stringr::str_trim(GenusOLD),
         Genus = case_when(
           stringr::str_detect(GenusOLD, "Pollachius") ~ "Gadus",
           CommonName == 'polar cod' ~ 'Arctogadus',
           CommonName == 'arctic cod' ~ 'Boreogadus',
           CommonName == 'pacific herring' ~ 'Clupea',
           CommonName == 'saffron cod' ~ 'Eleginus',
           CommonName == 'walleye pollock' ~ 'Gadus',
           CommonName == 'pacific cod' ~ 'Gadus',
           CommonName == 'flathead sole' ~ 'Hippoglossoides',
           CommonName == 'bering flounder' ~ 'Hippoglossoides',
           CommonName == 'yellowfin sole' ~ 'Limanda',
           CommonName == 'atka mackerel' ~ 'Pleurogrammus',
           CommonName == 'pacific sardine' ~ 'Sardinops',
           CommonName == 'pacific tomcod' ~ 'Microgadus',
           CommonName == 'sablefish' ~ 'Anoplopoma',
           CommonName == 'silver hake' ~ 'Merluccius',
           CommonName == 'steller sea lion' ~ 'Eumetopias',
           SpeciesNameOLD == 'Chionoecetes sp.' ~ 'Chionoecetes',
      TRUE ~ GenusOLD
    )
  )
 
Other_Entries %>%
  select(Genus) %>%
  table

NoGenusNames <- Other_Entries %>%
                   filter(is.na(Genus))

NoGenusNames %>% 
  select(CommonName)%>%
  table()
```

```{r StndrzFamily, eval=T, echo=T}
Other_Entries %>%
  select(FamilyOLD) %>%
  table()
  
NoFamilyNames <- Other_Entries %>%
                   filter(is.na(FamilyOLD))

NoFamilyNames %>% 
  select(CommonName)%>%
  table()

Other_Entries <- Other_Entries %>%
  mutate(FamilyOLD = as.character(FamilyOLD),
         FamilyOLD = stringr::str_trim(FamilyOLD),
         Family = case_when(
           CommonName == 'polar cod' ~ 'Gadidae',
           CommonName == 'arctic cod' ~ 'Gadidae',
           CommonName == 'pacific herring' ~ 'Clupeidae',
           CommonName == 'saffron cod' ~ 'Gadidae',
           CommonName == 'walleye pollock' ~ 'Gadidae',
           CommonName == 'pacific cod' ~ 'Gadidae',
           CommonName == 'flathead sole' ~ 'Pleuronectidae',
           CommonName == 'bering flounder' ~ 'Pleuronectidae',
           CommonName == 'yellowfin sole' ~ 'Pleuronectidae',
           CommonName == 'pacific tomcod' ~ 'Gadidae',
           CommonName == 'sablefish' ~ 'Anoplopoma',
           CommonName == 'silver hake' ~ 'Merluccius',
           CommonName == 'steller sea lion' ~ 'Eumetopias',
           Genus == 'Chionoecetes' ~ 'Oregoniidae',
      TRUE ~ FamilyOLD
    )
  )

#check
NoFamilyNames <- Other_Entries %>%
                   filter(is.na(Family))

NoFamilyNames %>% 
  select(CommonName)%>%
  table()
```


```{r StndrzSpecies, eval=T, echo=T}
Other_Entries %>%
  select(SpeciesNameOLD) %>%
  table
  
#It appears that in the Field_Values validation we don't
#want the binomial latin name as species, just the species 

Other_Entries <- Other_Entries %>%
  mutate(SpeciesNameOLD = as.character(SpeciesNameOLD),
         SpeciesNameOLD = stringr::str_trim(SpeciesNameOLD),
         Species = case_when(
           SpeciesNameOLD == 'Alopias sp.' ~ NA,
           SpeciesNameOLD == 'Anoplopoma fimbria' ~ 'fimbria',
           SpeciesNameOLD == 'Arctogadus glacialis' ~ 'glacialis',
           SpeciesNameOLD == 'Atheresthes evermanni' ~ 'evermanni',
           SpeciesNameOLD == 'Atheresthes stomias' ~ 'stomias',
           SpeciesNameOLD == 'Boreogadus saida' ~ 'saida',
           SpeciesNameOLD == 'Chionoecetes sp.' ~ NA,
           SpeciesNameOLD == 'Clupea pallasii' ~ 'pallasii',
           SpeciesNameOLD == 'Eleginus gracilis' ~ 'gracilis',
           SpeciesNameOLD == 'Gadus chalchogrammus' ~ 'chalcogrammus',  #fix misspelling
           SpeciesNameOLD == 'Gadus chalcogrammus' ~ 'chalcogrammus',
           SpeciesNameOLD == 'Gadus macrocephalus' ~ 'macrocephalus',
           SpeciesNameOLD == 'Hippoglossoides elassodon' ~ 'elassodon',
           SpeciesNameOLD == 'Hippoglossoides robustus' ~ 'robustus',
           SpeciesNameOLD == 'Ictalurus punctatus' ~ 'punctatus',
           SpeciesNameOLD == 'Lamna ditropis' ~ 'ditropis',
           SpeciesNameOLD == 'Limanda aspera' ~ 'aspera',
           SpeciesNameOLD == 'Microgadus proximus' ~ 'proximus',
           SpeciesNameOLD == 'Micropterus salmoides' ~ 'salmoides',
           SpeciesNameOLD == 'Notorynchus cepedianus' ~ 'cepedianus',
           SpeciesNameOLD == 'Paralithodes camtschaticus' ~ 'camtschaticus',
           SpeciesNameOLD == 'Pollachius pollachius' ~ 'chalcogrammus',  #this is walleye pollock, not pollock from the atlantic
           SpeciesNameOLD == 'Sardinops sagax' ~ 'sagax',
           SpeciesNameOLD == 'Somniosus ?' ~ NA,
           SpeciesNameOLD == 'Somniosus microcephalus' ~ 'microcephalus',
           SpeciesNameOLD == 'Somniosus pacificus' ~ 'pacificus')
  )

Other_Entries %>%
  select(Species) %>%
  table

NoSpeciesName <- Other_Entries %>%
  filter(is.na(Species))

NoSpeciesName %>%
  select(CommonName) %>%
  table()

Other_Entries <- Other_Entries %>%
  mutate(Species = case_when(
           CommonName== 'atka mackerel' ~ 'monopterygius',
           CommonName== 'pacific cod' ~ 'macrocephalus',
           CommonName== 'pacific herring' ~ 'pallasii',
           CommonName== 'pacific sardine' ~ 'sagax',
           CommonName== 'pacific tomcod' ~ 'proximus',
           CommonName== 'sablefish' ~ 'fimbria',
           CommonName== 'saffron cod' ~ 'gracilis',
           CommonName== 'silver hake' ~ 'bilinearis',
           CommonName== 'steller sea lion' ~ 'jubatus',
           CommonName== 'walleye pollock' ~ 'chalcogrammus',
           CommonName== 'thresher shark' ~ NA,
         TRUE ~ Species)
         )
           

Other_Entries %>%
  select(Species) %>%
  table

NoSpeciesName <- Other_Entries %>%
  filter(is.na(Species))
```

check common name, family, genus, species combos 
```{r}
Other_Entries %>%
  count(CommonName, Family, Genus, Species, name = "n") %>%
  arrange(Family, Genus, Species)
```


start a table to export and ask wes about
```{r}
question_entries1 <- Other_Entries %>%
  filter(CommonName %in% c("crab", "placeholder-unknown", "seal scat"))
question_entries2 <- Other_Entries %>%
  filter(is.na(CommonName)) %>%
  filter(is.na(Genus))
```



### Lab Fields ### 

if sample or DNA information is missing from the database, i don't think it's worth fixing those entries further so going to idenify those first... 
```{r}
#Preservation
Other_Entries %>%
  select(Preservation) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(Preservation = ifelse(Preservation=="---",NA,Preservation))%>%
  mutate(Preservation = recode(Preservation,
                               'dried' = 'Dried',
                               'Dry' = 'Dried',
                               'ethanol' = 'Ethanol',
                               'frozen' = 'Frozen')) #based on Validation keys in template

#SampleRoom 
Other_Entries %>%
  select(SampleRoom) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(SampleRoom = ifelse(SampleRoom=="---",NA,SampleRoom),
         SampleRoom = ifelse(SampleRoom=="NA",NA,SampleRoom),
         SampleRoom = ifelse(SampleRoom=="Ichthy","Ichthyology",SampleRoom),
         SampleRoom = ifelse(SampleRoom=="ichthyology","Ichthyology",SampleRoom),
         SampleRoom = ifelse(SampleRoom=="Icthology","Ichthyology",SampleRoom),
         SampleRoom = ifelse(SampleRoom=="Icthyology","Ichthyology",SampleRoom),
         SampleRoom = ifelse(SampleRoom=="ichthyology room","Ichthyology",SampleRoom),
         SampleRoom = ifelse(SampleRoom==213,"L213",SampleRoom))

#SampleLocation
Other_Entries %>%
  select(SampleLocation) %>%
  unlist()%>%
  table

 Other_Entries <- Other_Entries %>%
  mutate(SampleLocation = ifelse(SampleLocation=="---",NA,SampleLocation),
         SampleLocation = ifelse(SampleLocation=="gave back to charlie","returned",SampleLocation),
         SampleLocation = ifelse(SampleLocation=="juv pcod shelf","S14 Juvenile PCod",SampleLocation),
         SampleLocation = ifelse(SampleLocation=="juvenile cod shelf","S14 Juvenile PCod",SampleLocation),
         SampleLocation = ifelse(SampleLocation=="Red folder holder w/ age0 pcod","S14 Juvenile PCod",SampleLocation),
         SampleLocation = ifelse(SampleLocation=="P cod shelf","S15 PCod",SampleLocation),
         SampleLocation = ifelse(SampleLocation=="P. cod shelf","S15 PCod",SampleLocation),
         SampleLocation = ifelse(SampleLocation=="pcod shelf","S15 PCod",SampleLocation),
         SampleLocation = ifelse(SampleLocation=="Pcod Shelf","S15 PCod",SampleLocation))

#SampleSublocation 
Other_Entries %>%
  select(SampleSublocation) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(SampleSublocation = ifelse(SampleSublocation=="---",NA,SampleSublocation))

#ExtractedBy
Other_Entries %>%
  select(ExtractedBy) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(ExtractedBy = ifelse(ExtractedBy=="Juliana","Juliana Cornett",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="Jamie","Jamie Musbach",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="CS","Charlotte Springer",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="CT","Claire Tobin",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="JC","Juliana Cornett",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="Katie/Kirby","Katie DAmelio",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="KD","Katie DAmelio",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="KD/CS","Katie DAmelio",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="kirby","Kirby Karpan",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="Kirby","Kirby Karpan",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="Kirby/Dave N","Kirby Karpan",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="KL","Kim Ledger",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="HvN","Hanhvan Nguyen",ExtractedBy),                    
         ExtractedBy = ifelse(ExtractedBy=="N/A",NA,ExtractedBy))%>%
  mutate(ExtractedBy = ifelse(ExtractedBy=="---",NA,ExtractedBy))

#ExtractionDate
Other_Entries %>%
  select(ExtractionDate) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(ExtractionDate = ifelse(ExtractionDate=="---",NA,ExtractionDate),
         ExtractionDate = ifelse(ExtractionDate=="N/A",NA,ExtractionDate))%>%
  mutate(ExtractionDate = recode(ExtractionDate, #format YYYYMMDD
                                 '1/20/2021' = '20210120',
                                 '10/26/2022' = '20221026', 
                                 '11/17/2022' = '20221117',
                                 '12/1/2020' = '20201201', 
                                 '3/1/2023' = '20230301',
                                 '3/18/2021' = '20210318',
                                 '3/18/2021' = '20210318',
                                 '3/30/2021' = '20210330',
                                 '4/4/2023' = '20230404',
                                 '4/5/2023' = '20230405',
                                 '4/7/2021' = '20210407',
                                 '5/11/2022' = '20220511',
                                 '5/13/2022' = '20220513',
                                 '5/18/2022' = '20220518',
                                 '5/18/2023' = '20230518',
                                 '6/1/2023' = '20230601',
                                 '6/14/2023' = '20230614',
                                 '6/2/2023' = '20230602',
                                 '7/29/2022' = '20220729',
                                 '9/1/2022' = '20220901',
                                 '9/21/2022' = '20220921',
                                 '6/10/2022' = '20220610'))

#ExtractionMethod 
Other_Entries %>%
  select(ExtractionMethod) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(ExtractionMethod = recode(ExtractionMethod,
                                   'chelex' = 'Chelex',
                                 'QIAGEN DNeasy' = 'QIAGEN',
                                 'Macherey Nagel' = 'Macherey-Nagel',
                                 'Machery-Nagel' = 'Macherey-Nagel',
                                 'MN' = 'Macherey-Nagel'))%>%
  mutate(ExtractionMethod= ifelse(ExtractionMethod=="Unknown",NA,ExtractionMethod))


#ExtractionPlate
Other_Entries %>%
  select(ExtractionPlate) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(ExtractionPlate = ifelse(ExtractionPlate=="---",NA,ExtractionPlate),
         ExtractionPlate = ifelse(ExtractionPlate=="N/A",NA,ExtractionPlate))

Other_Entries <- Other_Entries %>%
  mutate(ExtractionPlate = ifelse(ExtractionPlate=="indiv tube","tube",ExtractionPlate),
         ExtractionPlate = ifelse(ExtractionPlate=="individual tube","tube",ExtractionPlate),
         ExtractionPlate = ifelse(ExtractionPlate=="IndividualTube","tube",ExtractionPlate),)

#ExtractionWell
Other_Entries %>%
  select(ExtractionWell) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(ExtractionWell = ifelse(ExtractionWell=="---",NA,ExtractionWell),
         ExtractionWell = ifelse(ExtractionWell=="N/A",NA,ExtractionWell))

#Need to decide if we want to have A01 or A1 for well locations and be consistent

#DNAngul
Other_Entries %>%
  select(DNAngul) %>%
  unlist()%>%
  table

Other_Entries %>% filter(is.na(DNAngul))
#0s could be valid here - but check if they are actually true 0's? lot's of NA's as well 

#DNAquality
Other_Entries %>%
  select(DNAquality) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(DNAquality = ifelse(DNAquality=="---",NA,DNAquality),
         DNAquality = ifelse(DNAquality=="n/a",NA,DNAquality))%>%
  mutate(DNAquality = str_trim(DNAquality))%>%
  mutate(DNAquality = recode(DNAquality,
                             #yes band and yes smear 
                             'HMW band and smear' = 'HMW bands and smearing',
                             'HMW band and smearing' = 'HMW bands and smearing',
                             'HMW Bands and smear' = 'HMW bands and smearing',
                             
                             #yes band and no smear
                             
                             #no band and yes smear 
                             'no  HMW bands and smearing' = 'no HMW bands and yes smearing',
                             'no HMW band and smearing' = 'no HMW bands and yes smearing', 
                             'No HMW bands and smearing' = 'no HMW bands and yes smearing',

                             #no band and no smear
                             'No HMW bands and no smearing' = 'no HMW bands and no smearing',
                             'No HMW bands and no smearing' = 'no HMW bands and no smearing',
                             'no HMW bands and no smearing (empty)' = 'no HMW bands and no smearing',
                             'no HMW bands and no smearing (empty) (empty)' = 'no HMW bands and no smearing',
                             'No HMW band' = 'no HMW bands and no smearing',
                             'No HMW Bands' = 'no HMW bands and no smearing'))


#DNAroom
Other_Entries %>%
  select(DNAroom) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(DNAroom = ifelse(DNAroom=="---",NA,DNAroom)) 

#DNAlocation
Other_Entries %>%
  select(DNAlocation) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(DNAlocation = ifelse(DNAlocation=="---",NA,DNAlocation)) 

#DNAsublocation 
Other_Entries %>%
  select(DNAsublocation) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(DNAsublocation = ifelse(DNAsublocation=="---",NA,DNAsublocation)) 


#WGS1
Other_Entries %>%
  select(WGSDateYYYYMMDD) %>%
  unlist()%>%
  table

Other_Entries %>%
  select(WGSBatchID) %>%
  unlist()%>%
  table

Other_Entries %>%
  select(WGSLaneID) %>%
  unlist()%>%
  table

#WGS2 - all blanks
Other_Entries %>%
  select(WGSDate2YYYYMMDD) %>%
  unlist()%>%
  table

Other_Entries %>%
  select(WGSBatchID2) %>%
  unlist()%>%
  table

Other_Entries %>%
  select(WGSLaneID2) %>%
  unlist()%>%
  table

#GTseq - all blanks 
Other_Entries %>%
  select(GTSeqDateYYYYMMDD) %>%
  unlist()%>%
  table

Other_Entries %>%
  select(GTSeqPanel) %>%
  unlist()%>%
  table

Other_Entries %>%
  select(GTSeqDate2YYYYMMDD) %>%
  unlist()%>%
  table

Other_Entries %>%
  select(GTSeqPanel2) %>%
  unlist()%>%
  table

Other_Entries %>%
  select(GTSeqDate3YYYYMMDD) %>%
  unlist()%>%
  table

Other_Entries %>%
  select(GTSeqPanel3) %>%
  unlist()%>%
  table

# Genetic ID - this column doesn't seem to exist

#Comments
Other_Entries %>%
  select(Comments) %>%
  unlist()%>%
  table
```


### Contact, Donating Lab, Collector

In the field descriptions 'Contact' is listed as a required field with 
instructions about field validations; however there are values listed in
the sheet. Additionally, contact is not listed in the 'ABLG-DB-working-copy'
sheet for upload. So perhaps that field has been replaced with 
'Collector' which has 69 unique values. 

```{r StndrzCollector, eval=T, echo=T}
Other_Entries %>%
  select(Collector) %>%
  table
  
Other_Entries <- Other_Entries %>%
  mutate(Collector = recode(Collector,
                            'E.Russ' = 'Elisa Russ',
                            'Lorenz Hauzer' = 'Lorenz Hauser',
                            'NMFS/ P.G./ S.K.' = 'NMFS/ S.K., P.G.'))%>%
  mutate(Collector = ifelse(Collector=="---",NA,Collector))

NoCollector <- Other_Entries %>%
  filter(is.na(Collector)) 
```

There are 12404 records without a collector recorded. Maybe that is fine provided
we have a contact/donating lab for the sample. 

Now take a stab at the Donating Lab

```{r StndrzCollector, eval=T, echo=T}
Other_Entries %>%
  select(DonatingLab) %>%
  table
  
Other_Entries <- Other_Entries %>%
  mutate(DonatingLab = recode(DonatingLab,
                              'ABL freezer' = 'ABL Freezer',
                              'ADFG Anchorage' = 'ADFG-Anchorage',
                              'Canino_Spies' = 'Canino/Spies',
                            'LPW/Waters' = 'AFSC-LPW',
                            'Matt Craig' = 'Matt Craig SWFSC',
                            'Spies' = 'Ingrid Spies'))%>%
  mutate(DonatingLab = ifelse(DonatingLab=="---",NA,DonatingLab))

sum(is.na(Other_Entries$DonatingLab))
```

```{r UnknwnRcd, eval=T, echo=T}
question_entries3 <- Other_Entries %>%
  filter(is.na(DonatingLab)&is.na(Collector))

question_entries4 <- Other_Entries %>%
  filter(Collector == "Courtney?")
```

What are the samples from 

sharks, sablefish, seal scat, pcod


#### Collection Year, Month, Day

```{r Year, eval=T, echo=F}
Other_Entries %>%
  select(CollectionYear) %>%
  unlist()%>%
  table
```

```{r Month, eval=T, echo=F}
Other_Entries %>%
  select(CollectionMonth) %>%
  unlist()%>%
  table
```

```{r Year, eval=T, echo=F}
Other_Entries %>%
  select(CollectionDay) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(CollectionDay = recode(CollectionDay,
                                '1' = '01',
                                '2' = '02',
                                '3' = '03',
                                '4' = '04',
                                '5' = '05',
                                '6' = '06',
                                '7' = '07',
                                '8' = '08',
                                '9' = '09')) %>%
  mutate(CollectionDay = ifelse(CollectionDay == "?", NA, CollectionDay))
```


### Location

```{r MR, eval=T, echo=F}
Other_Entries %>%
  select(MarineRegion) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(MarineRegion = ifelse(MarineRegion=="---",NA,MarineRegion),
         MarineRegion = ifelse(MarineRegion=="AI","Aleutian Islands",MarineRegion),
         MarineRegion = ifelse(MarineRegion=="CBS","Coastal BC",MarineRegion),
         MarineRegion = ifelse(MarineRegion=="WGOA","wGOA",MarineRegion),
         MarineRegion = ifelse(MarineRegion=="stat area 505932",505932,MarineRegion),
         MarineRegion = ifelse(MarineRegion=="Eastern GOA","eGOA",MarineRegion),
         MarineRegion = ifelse(MarineRegion=="PWS","Prince William Sound",MarineRegion),)

##change the stat areas to be region?
Other_Entries <- Other_Entries %>%
  mutate(Location1 = ifelse(MarineRegion %in% c("476004/476003", "485935", "505932", "515905", "515936"), MarineRegion, Location1))
 
Other_Entries <- Other_Entries %>%
  mutate(MarineRegion = ifelse(MarineRegion=="476004/476003","Prince William Sound",MarineRegion),
         MarineRegion = ifelse(MarineRegion=="485935","Prince William Sound",MarineRegion),
         MarineRegion = ifelse(MarineRegion=="505932","Cook Inlet",MarineRegion),
         MarineRegion = ifelse(MarineRegion=="515905","Cook Inlet",MarineRegion),
         MarineRegion = ifelse(MarineRegion=="515936","Cook Inlet",MarineRegion))
  

Other_Entries %>%
  select(StateCountry) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(StateCountry = ifelse(StateCountry=="---",NA,StateCountry),
         StateCountry = ifelse(StateCountry=="AK","Alaska, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="AK, USA","Alaska, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="AK; USA","Alaska, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="AK;USA","Alaska, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="Alaska","Alaska, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="Alaska USA","Alaska, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="BC, CAN","British Columbia, Cananda",StateCountry),
         StateCountry = ifelse(StateCountry=="BC;Canada","British Columbia, Cananda",StateCountry),
         StateCountry = ifelse(StateCountry=="CA, USA","California, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="CA;USA","California, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="CTl; USA","Connecticut, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="Eureka, CA","Eureka, California, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="Gaviota","Gaviota, California, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="Hokkaido;JP","Hokkaido, Japan",StateCountry),
         StateCountry = ifelse(StateCountry=="KOREA","Korea",StateCountry),
         StateCountry = ifelse(StateCountry=="Magdalena Bay, MX","Magdalena Bay, Mexico",StateCountry),
         StateCountry = ifelse(StateCountry=="MX","Mexico",StateCountry),
         StateCountry = ifelse(StateCountry=="OR, USA","Oregon, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="OR;USA","Oregon, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="Pescadero, CA","Pescadero, California, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="Punta Colonet, BC, MX","Punta Colonet, Baja California, Mexico",StateCountry),
         StateCountry = ifelse(StateCountry=="RI;USA","Rhode Island, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="Santa Cruz, CA","Santa Cruz, California, USA",StateCountry),
         StateCountry = ifelse(StateCountry=="WA;USA","Washington, USA",StateCountry))
```

```{r FWWS, eval=T, echo=F}
Other_Entries %>%
  select(FreshwaterWatershed) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(FreshwaterWatershed = ifelse(FreshwaterWatershed=="---",NA,FreshwaterWatershed))
```

```{r Loc1, eval=T, echo=F}
Other_Entries %>%
  select(Location1) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(Location1 = ifelse(Location1=="---",NA,Location1))
```

```{r Loc2, eval=T, echo=F}
Other_Entries %>%
  select(Location2) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(Location2 = ifelse(Location2=="---",NA,Location2))
```


```{r Lat, eval=T, echo=F}
Other_Entries %>%
  select(StartLatitudeDD) %>%
  unlist()%>%
  table

#0 isn't valid
Other_Entries <- Other_Entries %>%
  mutate(StartLatitudeDD = ifelse(StartLatitudeDD==0,NA,StartLatitudeDD))

Other_Entries %>%
  select(EndLatitudeDD) %>%
  unlist()%>%
  table
#no entries

Other_Entries <- Other_Entries %>%
  mutate(EndLatitudeDD = ifelse(EndLatitudeDD=="---",NA,EndLatitudeDD))
```

```{r Long, eval=T, echo=F}
Other_Entries %>%
  select(StartLongitudeDD) %>%
  unlist()%>%
  table

Other_Entries %>%
  select(EndLongitudeDD) %>%
  unlist()%>%
  table
#no entries

Other_Entries <- Other_Entries %>%
  mutate(EndLongitudeDD = ifelse(EndLongitudeDD=="---",NA,EndLongitudeDD))
```

```{r Long, eval=T, echo=F}
Other_Entries %>%
  select(EstimatedLatLongYesNo) %>%
  unlist()%>%
  table

#no entries
```

The estimated or not estimated lat/long for each collection should be 
entered with the metadata... not entered for any entry 

### Survey Specific Fields maybe not super relevant to salmon

```{r Survey, eval=T, echo=F}
Other_Entries %>%
  select(CruiseID) %>%
  unlist()%>%
  table

Other_Entries %>%
  select(HaulID) %>%
  unlist()%>%
  table

Other_Entries %>%
  select(StationID) %>%
  unlist()%>%
  table

Other_Entries %>%
  select(Equipment_Gear) %>%
  unlist()%>%
  table

Other_Entries %>%
  select(DepthM) %>%
  unlist()%>%
  table
#is 0 valid for depth in meters?

Other_Entries %>%
  select(StartDepthM) %>%
  unlist()%>%
  table
#nothing

Other_Entries %>%
  select(EndDepthM) %>%
  unlist()%>%
  table
#nothing

Other_Entries %>%
  select(TempC) %>%
  unlist()%>%
  table
## are the 0's real data??? i would think not...

Other_Entries %>%
  select(Habitat) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(CruiseID = ifelse(CruiseID=="---",NA,CruiseID))%>%
  mutate(HaulID = ifelse(HaulID=="---",NA,HaulID))%>%
   mutate(StationID = ifelse(StationID=="---",NA,StationID))%>%
  mutate(Habitat = ifelse(Habitat=="---",NA,Habitat))%>%
   mutate(Equipment_Gear = ifelse(Equipment_Gear=="---",NA,Equipment_Gear))%>%
  mutate(DepthM = ifelse(DepthM==0,NA,DepthM))%>%
  mutate(TempC = ifelse(TempC==0,NA,TempC))
```

#### Morphometrics

```{r Survey, eval=T, echo=F}
Other_Entries %>%
  select(LengthMM) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(LengthMM = ifelse(LengthMM==0,NA,LengthMM))

Other_Entries %>%
  select(LengthType) %>%
  unlist()%>%
  table
#empty

Other_Entries %>%
  select(FishDepthMM) %>%
  unlist()%>%
  table
#empty

Other_Entries %>%
  select(WeightG) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(WeightG = ifelse(WeightG==0,NA,WeightG))


Other_Entries %>%
  select(Lifestage) %>%
  unlist()%>%
  table

# should this be more of a categorical lifestage not age? smolt, juvenile, adult etc? We have age in years
Other_Entries <- Other_Entries %>%
  mutate(Lifestage = ifelse(Lifestage=="---",NA,Lifestage),
         Lifestage = ifelse(Lifestage=="Adult","adult",Lifestage),
         Lifestage = ifelse(Lifestage=="Juvenile","juvenile",Lifestage),
         Lifestage = ifelse(Lifestage=="Juveniles","juvenile",Lifestage))

Other_Entries %>%
  filter(Lifestage == 7)    #####these are pcod ages????

Other_Entries <- Other_Entries %>%
  mutate(AgeYRS = ifelse(Lifestage %in% c(1,2,4,5,6,7,9), Lifestage, AgeYRS),
         Lifestage = ifelse(Lifestage %in% c(1,2,4,5,6,7,9), NA, Lifestage))

Other_Entries %>%
  select(AgeYRS) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(AgeYRS = ifelse(AgeYRS=="---",NA,AgeYRS),
         AgeYRS = ifelse(AgeYRS==0,NA,AgeYRS),
         AgeYRS = ifelse(AgeYRS=="16.100000000000001",16.1,AgeYRS),
         AgeYRS = ifelse(AgeYRS=="17.100000000000001",17.1,AgeYRS),
         AgeYRS = ifelse(AgeYRS=="18.100000000000001",18.1,AgeYRS),
         AgeYRS = ifelse(AgeYRS=="18.399999999999999",18.4,AgeYRS),
         AgeYRS = ifelse(AgeYRS=="19.100000000000001",19.1,AgeYRS),
         AgeYRS = ifelse(AgeYRS=="20.100000000000001",20.1,AgeYRS))

Other_Entries %>%
  select(Sex) %>%
  unlist()%>%
  table

Other_Entries %>%
  filter(Sex %in% c(1,2,3))

### numbers for sex represent what? 
Other_Entries <- Other_Entries %>%
  mutate(Sex = ifelse(Sex=="---",NA,Sex),
         Sex = ifelse(Sex=="NA",NA,Sex),
         Sex = ifelse(Sex=="F","female",Sex),
         Sex = ifelse(Sex=="M","male",Sex),
         Sex = ifelse(Sex=="U","unsexed",Sex))

Other_Entries <- Other_Entries %>%
  mutate(Sex = ifelse(Sex == 1 & CommonName == "pacific cod", NA, Sex),
         Sex = ifelse(Sex == 2 & CommonName == "pacific cod", NA, Sex),
         Sex = ifelse(Sex == 3 & CommonName == "pacific cod", NA, Sex),
         Sex = ifelse(Sex == 1 & CommonName == "yellowfin sole", "male", Sex),
         Sex = ifelse(Sex == 2 & CommonName == "yellowfin sole", "female", Sex),
         Sex = ifelse(Sex == 3 & CommonName == "yellowfin sole", "unsexed", Sex),
         Sex = ifelse(Sex == 1 & CommonName == "bering flounder", "male", Sex),
         Sex = ifelse(Sex == 2 & CommonName == "bering flounder", "female", Sex),
         Sex = ifelse(Sex == 3 & CommonName == "bering flounder", "unsexed", Sex))
  
Other_Entries %>%
  select(CaptureTime) %>%
  unlist()%>%
  table
#empty

Other_Entries %>%
  select(ReleaseTime) %>%
  unlist()%>%
  table
#empty

Other_Entries %>%
  select(IdentifiedBy) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(IdentifiedBy = ifelse(IdentifiedBy=="---",NA,IdentifiedBy))
```

any samples there is no sample or dna location info?
```{r}
question_entries5 <- Other_Entries %>%
  filter(is.na(SampleRoom) & is.na(SampleLocation) & is.na(SampleSublocation) & is.na(DNAroom) & is.na(DNAlocation) & is.na(DNAsublocation) & is.na(ExtractionPlate))

question_entries5 %>%
  count(CommonName, Family, Genus, Species, DonatingLab, Collector, name = "n") %>%
  arrange(desc(n))
```



```{r}
to_check <- question_entries1 %>% mutate(reason = "no ID") %>%
  bind_rows(question_entries2 %>% mutate(reason = "no ID")) %>%
  bind_rows(question_entries3 %>% mutate(reason = "no DonatingLab or Collector")) %>%
  bind_rows(question_entries4 %>% mutate(reason = "no DonatingLab or Collector"))

#write.csv(to_check, "missing_info.csv")
```


### adding ID/Lab/location information based info from Wes

```{r}
Other_Entries <- Other_Entries %>%
  filter(ABLG != 12351) %>%    #blank placeholder entry
  mutate(Genus = ifelse(ABLG == 31096, "Sebastes", Genus)) %>%
  mutate(DonatingLab = ifelse(ABLG >= 10208 & ABLG <= 10407, "ADFG", DonatingLab)) %>% 
  mutate(DonatingLab = ifelse(ABLG >= 38695 & ABLG <= 38753, "Susanne McDermott", DonatingLab)) %>%
  mutate(DNAroom = ifelse(DonatingLab == "Kim Ledger", "L212", DNAroom),
         DNAlocation = ifelse(DonatingLab == "Kim Ledger", "F212_1", DNAroom),
         DonatingLab = ifelse(DonatingLab == "Kim Ledger", "UW_fish_coll", DonatingLab))
```



welp, the larval gadids from Johanna that were genetically ID'd never had their info updated in the database.  
i have that info so will do that here... 
```{r}
gadid_2024 <- read.csv("ABL_dat/2024_gadid_for_genetics_w_results.csv") 

pcod_ids <- gadid_2024$ABLG[gadid_2024$Genetic_ID == "Gadus macrocephalus"]
pollock_ids <- gadid_2024$ABLG[gadid_2024$Genetic_ID == "Gadus chalcogrammus"]
saf_ids <- gadid_2024$ABLG[gadid_2024$Genetic_ID == "Eleginus gracilis"]
na_ids <- gadid_2024$ABLG[is.na(gadid_2024$Genetic_ID)]

Other_Entries <- Other_Entries %>%
  mutate(
    CommonName = if_else(ABLG %in% pcod_ids, "pacific cod", CommonName),
    Genus = if_else(ABLG %in% pcod_ids, "Gadus", Genus),
    Family = if_else(ABLG %in% pcod_ids, "Gadidae", Family),
    Species = if_else(ABLG %in% pcod_ids, "macrocephalus", Species),
    CommonName = if_else(ABLG %in% pollock_ids, "walleye pollock", CommonName),
    Genus = if_else(ABLG %in% pollock_ids, "Gadus", Genus),
    Family = if_else(ABLG %in% pollock_ids, "Gadidae", Family),
    Species = if_else(ABLG %in% pollock_ids, "chalcogrammus", Species),
    CommonName = if_else(ABLG %in% saf_ids, "saffron cod", CommonName),
    Genus = if_else(ABLG %in% saf_ids, "Eleginus", Genus),
    Family = if_else(ABLG %in% saf_ids, "Gadidae", Family),
    Species = if_else(ABLG %in% saf_ids, "gracilis", Species),
    CommonName = if_else(ABLG %in% na_ids, NA, CommonName),
    Genus = if_else(ABLG %in% na_ids, NA, Genus),
    Family = if_else(ABLG %in% na_ids, "Gadidae", Family),
    Species = if_else(ABLG %in% na_ids, NA, Species))
```

