---
title: "Building a curated Alaska Fisheries Science Center genetic sample repository to facilitate long-term use of molecular information for fisheries management"
author: ""
date: "11/29/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Summary
The Alaska Fisheries Science Center Genetics Program (AFSC GP) conducts 
research on Pacific salmon, groundfish, and forage fish species to comply 
with regional, national, and international agreements and treaties for 
managing marine resources. Much of this work relies on the collection of
samples, both fishery dependent and independent, to evaluate the existence 
and extent of population genetic structure in a variety of species. 
Currently, the genetics program manages a database of these samples, associated 
collection information, and laboratory data in FileMaker software. Filemaker is 
a relational database application with a graphical user interface. Recent 
limitations with the current database highlight the need for better records 
management. This project's primary goal is to work with the Alaska Fisheries 
Information Network (AKFIN) to develop a database model and
repository for AFSC GP's inventory of genetic samples. Additionally, a data 
management and upload tool in Oracle APEX will be created to interface 
with these data. These tables will be used for storing sample information 
including species, tissue type, collection information, and laboratory processing
data. Effective organization of these samples will aid in both the evaluation 
of population genetic trends across a diverse array of taxa as well as 
the development of molecular resources for fisheries
management. This project would bring consistency to AFSC GP's records 
management, as current salmon bycatch analyses rely heavily on AKFIN's database.

## Description of the current database
```{r RdInDat, eval=F, echo = T}
FMdat <- readxl::read_xlsx("./ABL_dat/20260127_export.xlsx")

FMupload <- readxl::read_xlsx("./ABL_dat/upload_template_fieldValidation.xlsx",
                              sheet='Field_Descriptions',
                              range='A1:B67',
                              col_names=F)
```

```{r FieldDescip, eval=T, echo=F}
FMupload %>%
  rename('Field' = 1,
         'Description' = 2) %>%
  knitr::kable()%>%
  kableExtra::kable_styling(full_width = F)

```

The current Filemaker database export (20260127) has a total of 68 fields.

Pat will take care of the salmon and Diana will address the rockfish.

```{r Salmon, eval=F, echo = T}
Salmon_Entries <- FMdat %>%
  filter(grepl(pattern='Salmonidae',FamilyOLD) | grepl(pattern='King|king|Chinook|chinook|Pink|pink|Coho|coho|Chum|chum|salmon|Steelhead|steelhead',CommonNameOLD)|grepl(pattern='Oncorhynchus',GenusOLD)|grepl(pattern='Oncorhynchus',SpeciesNameOLD)) %>%
  filter(!SpeciesNameOLD %in% c('Lamna ditropis', 'Paralithodes camtschaticus'))

Rockfish_Entries <- FMdat %>%
  filter(
    grepl(pattern='Sebastidae',FamilyOLD) |
      grepl(pattern='Rockfish|rockfish',CommonNameOLD)|
      grepl(pattern='Sebastes|SEBASTES|Sebastolobus|Sebastulobus',GenusOLD)|
      grepl(pattern='Sebastes|Sebastolobus',SpeciesNameOLD))

Other_Entries <- FMdat %>%
  anti_join(Salmon_Entries) %>%
  anti_join(Rockfish_Entries)
```


## Not Salmoniformes or Sebastes/Sebastolobus Entries

check what's there 
```{r Genera, eval=T, echo=T}
Other_Entries %>% 
  select(GenusOLD) %>%
  table()
```

```{r}
Other_Entries %>% 
  select(CommonNameOLD)%>%
  table()
```

```{r}
Other_Entries %>% 
  select(SpeciesNameOLD)%>%
  table()
```


#### Common Names

I'm going to follow Pat and first standardize the common names.
```{r StndrzCN, eval=T, echo=T}
Other_Entries <- Other_Entries %>%
  mutate(CommonNameOLD = as.character(CommonNameOLD),
         CommonNameOLD = stringr::str_trim(CommonNameOLD),
         CommonNameOLD = stringr::str_to_lower(CommonNameOLD),
         CommonName = case_when(
           stringr::str_detect(CommonNameOLD, "pollock") ~ "walleye pollock",
           stringr::str_detect(CommonNameOLD, "pacific cod") ~ "pacific cod",
           stringr::str_detect(CommonNameOLD, "channel catfish") ~ "channel catfish",
           stringr::str_detect(CommonNameOLD, "largemouth bass") ~ "largemouth bass",
           stringr::str_detect(CommonNameOLD, "pacific sardine") ~ "pacific sardine",
           stringr::str_detect(CommonNameOLD, "steller sea lion") ~ "steller sea lion",
      # Default: keep original
      TRUE ~ CommonNameOLD
    )
  )

#what are 'crab', 'placeholder-unknown', 'seal scat', 'sleeper shark ?'
Other_Entries %>% filter(CommonNameOLD == 'sleeper shark ?')

Other_Entries %>% 
  select(CommonName)%>%
  table()

Other_Entries %>% nrow() == (Other_Entries %>% 
                                select(CommonName)%>%
                                  table() %>% sum())
```

Dang, some of the common names are absent. Hopefully there is an entry in 
the genus and species. 

```{r NoCN, eval=T, echo=T}
NoCommonNames <- Other_Entries %>%
                   filter(is.na(CommonName))

NoCommonNames %>% 
  select(SpeciesNameOLD)%>%
  table()
```

add common name for those that we can 
```{r}
Other_Entries <- Other_Entries %>%
  mutate(
    CommonName = case_when(
      SpeciesNameOLD == 'Arctogadus glacialis' ~ 'polar cod',
      SpeciesNameOLD == 'Boreogadus saida' ~ 'arctic cod',
      SpeciesNameOLD == 'Clupea pallasii' ~ 'pacific herring',
      SpeciesNameOLD == 'Eleginus gracilis' ~ 'saffron cod',
      SpeciesNameOLD == 'Gadus chalchogrammus' ~ 'walleye pollock',
      SpeciesNameOLD == 'Gadus chalcogrammus' ~ 'walleye pollock',
      SpeciesNameOLD == 'Gadus macrocephalus' ~ 'pacific cod',
      SpeciesNameOLD == 'Hippoglossoides elassodon' ~ 'flathead sole',
      SpeciesNameOLD == 'Hippoglossoides robustus' ~ 'bering flounder',
      SpeciesNameOLD == 'Limanda aspera' ~ 'yellowfin sole',
      SpeciesNameOLD == 'Microgadus proximus' ~ 'pacific tomcod',
      TRUE ~ CommonName
    )
  )
```

```{r}
NoCommonNames <- Other_Entries %>%
                   filter(is.na(CommonName))

NoCommonNames %>% 
  select(SpeciesNameOLD)%>%
  table()
```

also some missing ID's for juvenile cod from Johanna here.


### Genus, Family, and Species
Now that *most* the common names are filled in we can work with the genus name.

```{r StndrzGenus, eval=T, echo=T}
Other_Entries %>%
  select(GenusOLD) %>%
  table
#incomplete and incorrect in some cases (Pollachius)  

Other_Entries <- Other_Entries %>%
  mutate(GenusOLD = as.character(GenusOLD),
         GenusOLD = stringr::str_trim(GenusOLD),
         Genus = case_when(
           stringr::str_detect(GenusOLD, "Pollachius") ~ "Gadus",
           CommonName == 'polar cod' ~ 'Arctogadus',
           CommonName == 'arctic cod' ~ 'Boreogadus',
           CommonName == 'pacific herring' ~ 'Clupea',
           CommonName == 'saffron cod' ~ 'Eleginus',
           CommonName == 'walleye pollock' ~ 'Gadus',
           CommonName == 'pacific cod' ~ 'Gadus',
           CommonName == 'flathead sole' ~ 'Hippoglossoides',
           CommonName == 'bering flounder' ~ 'Hippoglossoides',
           CommonName == 'yellowfin sole' ~ 'Limanda',
           CommonName == 'atka mackerel' ~ 'Pleurogrammus',
           CommonName == 'pacific sardine' ~ 'Sardinops',
           CommonName == 'pacific tomcod' ~ 'Microgadus',
           CommonName == 'sablefish' ~ 'Anoplopoma',
           CommonName == 'silver hake' ~ 'Merluccius',
           CommonName == 'steller sea lion' ~ 'Eumetopias',
           SpeciesNameOLD == 'Chionoecetes sp.' ~ 'Chionoecetes',
      TRUE ~ GenusOLD
    )
  )
 
Other_Entries %>%
  select(Genus) %>%
  table

NoGenusNames <- Other_Entries %>%
                   filter(is.na(Genus))

NoGenusNames %>% 
  select(CommonName)%>%
  table()
```

```{r StndrzFamily, eval=T, echo=T}
Other_Entries %>%
  select(FamilyOLD) %>%
  table()
  
NoFamilyNames <- Other_Entries %>%
                   filter(is.na(FamilyOLD))

NoFamilyNames %>% 
  select(CommonName)%>%
  table()

Other_Entries <- Other_Entries %>%
  mutate(FamilyOLD = as.character(FamilyOLD),
         FamilyOLD = stringr::str_trim(FamilyOLD),
         Family = case_when(
           CommonName == 'polar cod' ~ 'Gadidae',
           CommonName == 'arctic cod' ~ 'Gadidae',
           CommonName == 'pacific herring' ~ 'Clupeidae',
           CommonName == 'saffron cod' ~ 'Gadidae',
           CommonName == 'walleye pollock' ~ 'Gadidae',
           CommonName == 'pacific cod' ~ 'Gadidae',
           CommonName == 'flathead sole' ~ 'Pleuronectidae',
           CommonName == 'bering flounder' ~ 'Pleuronectidae',
           CommonName == 'yellowfin sole' ~ 'Pleuronectidae',
           CommonName == 'pacific tomcod' ~ 'Gadidae',
           CommonName == 'sablefish' ~ 'Anoplopoma',
           CommonName == 'silver hake' ~ 'Merluccius',
           CommonName == 'steller sea lion' ~ 'Eumetopias',
           Genus == 'Chionoecetes' ~ 'Oregoniidae',
      TRUE ~ FamilyOLD
    )
  )

#check
NoFamilyNames <- Other_Entries %>%
                   filter(is.na(Family))

NoFamilyNames %>% 
  select(CommonName)%>%
  table()
```


```{r StndrzSpecies, eval=T, echo=T}
Other_Entries %>%
  select(SpeciesNameOLD) %>%
  table
  
#It appears that in the Field_Values validation we don't
#want the binomial latin name as species, just the species 

Other_Entries <- Other_Entries %>%
  mutate(SpeciesNameOLD = as.character(SpeciesNameOLD),
         SpeciesNameOLD = stringr::str_trim(SpeciesNameOLD),
         Species = case_when(
           SpeciesNameOLD == 'Alopias sp.' ~ NA,
           SpeciesNameOLD == 'Anoplopoma fimbria' ~ 'fimbria',
           SpeciesNameOLD == 'Arctogadus glacialis' ~ 'glacialis',
           SpeciesNameOLD == 'Atheresthes evermanni' ~ 'evermanni',
           SpeciesNameOLD == 'Atheresthes stomias' ~ 'stomias',
           SpeciesNameOLD == 'Boreogadus saida' ~ 'saida',
           SpeciesNameOLD == 'Chionoecetes sp.' ~ NA,
           SpeciesNameOLD == 'Clupea pallasii' ~ 'pallasii',
           SpeciesNameOLD == 'Eleginus gracilis' ~ 'gracilis',
           SpeciesNameOLD == 'Gadus chalchogrammus' ~ 'chalcogrammus',  #fix misspelling
           SpeciesNameOLD == 'Gadus chalcogrammus' ~ 'chalcogrammus',
           SpeciesNameOLD == 'Gadus macrocephalus' ~ 'macrocephalus',
           SpeciesNameOLD == 'Hippoglossoides elassodon' ~ 'elassodon',
           SpeciesNameOLD == 'Hippoglossoides robustus' ~ 'robustus',
           SpeciesNameOLD == 'Ictalurus punctatus' ~ 'punctatus',
           SpeciesNameOLD == 'Lamna ditropis' ~ 'ditropis',
           SpeciesNameOLD == 'Limanda aspera' ~ 'aspera',
           SpeciesNameOLD == 'Microgadus proximus' ~ 'proximus',
           SpeciesNameOLD == 'Micropterus salmoides' ~ 'salmoides',
           SpeciesNameOLD == 'Notorynchus cepedianus' ~ 'cepedianus',
           SpeciesNameOLD == 'Paralithodes camtschaticus' ~ 'camtschaticus',
           SpeciesNameOLD == 'Pollachius pollachius' ~ 'chalcogrammus',  #this is walleye pollock, not pollock from the atlantic
           SpeciesNameOLD == 'Sardinops sagax' ~ 'sagax',
           SpeciesNameOLD == 'Somniosus ?' ~ NA,
           SpeciesNameOLD == 'Somniosus microcephalus' ~ 'microcephalus',
           SpeciesNameOLD == 'Somniosus pacificus' ~ 'pacificus')
  )

Other_Entries %>%
  select(Species) %>%
  table

NoSpeciesName <- Other_Entries %>%
  filter(is.na(Species))

NoSpeciesName %>%
  select(CommonName) %>%
  table()

Other_Entries <- Other_Entries %>%
  mutate(Species = case_when(
           CommonName== 'atka mackerel' ~ 'monopterygius',
           CommonName== 'pacific cod' ~ 'macrocephalus',
           CommonName== 'pacific herring' ~ 'pallasii',
           CommonName== 'pacific sardine' ~ 'sagax',
           CommonName== 'pacific tomcod' ~ 'proximus',
           CommonName== 'sablefish' ~ 'fimbria',
           CommonName== 'saffron cod' ~ 'gracilis',
           CommonName== 'silver hake' ~ 'bilinearis',
           CommonName== 'steller sea lion' ~ 'jubatus',
           CommonName== 'walleye pollock' ~ 'chalcogrammus',
           CommonName== 'thresher shark' ~ NA,
         TRUE ~ Species)
         )
           

Other_Entries %>%
  select(Species) %>%
  table

NoSpeciesName <- Other_Entries %>%
  filter(is.na(Species))
```

check common name, family, genus, species combos 
```{r}
Other_Entries %>%
  count(CommonName, Family, Genus, Species, name = "n") %>%
  arrange(Family, Genus, Species)
```

looks better - still need to consult wes about what some of these entries are.... 

### Lab Fields ### 

if sample or DNA information is missing from the database, i don't think it's worth fixing those entries further so going to idenify those first... 
```{r}
#Preservation
Other_Entries %>%
  select(Preservation) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(Preservation = ifelse(Preservation=="---",NA,Preservation))%>%
  mutate(Preservation = recode(Preservation,
                               'dried' = 'Dried',
                               'Dry' = 'Dried',
                               'ethanol' = 'Ethanol',
                               'frozen' = 'Frozen')) #based on Validation keys in template

#DNAroom
Other_Entries %>%
  select(DNAroom) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(DNAroom = ifelse(DNAroom=="---",NA,DNAroom)) 

#DNAlocation
Other_Entries %>%
  select(DNAlocation) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(DNAlocation = ifelse(DNAlocation=="---",NA,DNAlocation)) 

#DNAsublocation 
Other_Entries %>%
  select(DNAsublocation) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(DNAsublocation = ifelse(DNAsublocation=="---",NA,DNAsublocation)) 

#ExtractedBy
Other_Entries %>%
  select(ExtractedBy) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(ExtractedBy = ifelse(ExtractedBy=="Juliana","Juliana Cornett",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="Jamie","Jamie Musbach",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="CS","Charlotte Springer",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="CT","Claire Tobin",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="JC","Juliana Cornett",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="Katie/Kirby","Katie DAmelio",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="KD","Katie DAmelio",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="KD/CS","Katie DAmelio",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="kirby","Kirby Karpan",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="Kirby","Kirby Karpan",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="Kirby/Dave N","Kirby Karpan",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="KL","Kim Ledger",ExtractedBy),
         ExtractedBy = ifelse(ExtractedBy=="N/A",NA,ExtractedBy))%>%
  mutate(ExtractedBy = ifelse(ExtractedBy=="---",NA,ExtractedBy))

#ExtractionDate
Other_Entries %>%
  select(ExtractionDate) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(ExtractionDate = ifelse(ExtractionDate=="---",NA,ExtractionDate),
         ExtractionDate = ifelse(ExtractionDate=="N/A",NA,ExtractionDate))%>%
  mutate(ExtractionDate = recode(ExtractionDate, #format YYYYMMDD
                                 '1/20/2021' = '20210120',
                                 '10/26/2022' = '20221026', 
                                 '11/17/2022' = '20221117',
                                 '12/1/2020' = '20201201', 
                                 '3/1/2023' = '20230301',
                                 '3/18/2021' = '20210318',
                                 '3/18/2021' = '20210318',
                                 '3/30/2021' = '20210330',
                                 '4/4/2023' = '20230404',
                                 '4/5/2023' = '20230405',
                                 '4/7/2021' = '20210407',
                                 '5/11/2022' = '20220511',
                                 '5/13/2022' = '20220513',
                                 '5/18/2022' = '20220518',
                                 '5/18/2023' = '20230518',
                                 '6/1/2023' = '20230601',
                                 '6/14/2023' = '20230614',
                                 '6/2/2023' = '20230602',
                                 '7/29/2022' = '20220729',
                                 '9/1/2022' = '20220901',
                                 '9/21/2022' = '20220921',
                                 '6/10/2022' = '20220610'))

#ExtractionMethod 
Other_Entries %>%
  select(ExtractionMethod) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(ExtractionMethod = recode(ExtractionMethod,
                                   'chelex' = 'Chelex',
                                 'QIAGEN DNeasy' = 'QIAGEN',
                                 'Macherey Nagel' = 'Macherey-Nagel',
                                 'Machery-Nagel' = 'Macherey-Nagel',
                                 'MN' = 'Macherey-Nagel'))%>%
  mutate(ExtractionMethod= ifelse(ExtractionMethod=="Unknown",NA,ExtractionMethod))


#ExtractionPlate
Other_Entries %>%
  select(ExtractionPlate) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(ExtractionPlate = ifelse(ExtractionPlate=="---",NA,ExtractionPlate),
         ExtractionPlate = ifelse(ExtractionPlate=="N/A",NA,ExtractionPlate))

Other_Entries <- Other_Entries %>%
  mutate(ExtractionPlate = ifelse(ExtractionPlate=="indiv tube","tube",ExtractionPlate),
         ExtractionPlate = ifelse(ExtractionPlate=="individual tube","tube",ExtractionPlate),
         ExtractionPlate = ifelse(ExtractionPlate=="IndividualTube","tube",ExtractionPlate),)

#ExtractionWell
Other_Entries %>%
  select(ExtractionWell) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(ExtractionWell = ifelse(ExtractionWell=="---",NA,ExtractionWell),
         ExtractionWell = ifelse(ExtractionWell=="N/A",NA,ExtractionWell))

#Need to decide if we want to have A01 or A1 for well locations and be consistent

#DNAngul
Other_Entries %>%
  select(DNAngul) %>%
  unlist()%>%
  table

Other_Entries %>% filter(is.na(DNAngul))
#0s could be valid here - but check if they are actually true 0's? lot's of NA's as well 

#DNAquality
Other_Entries %>%
  select(DNAquality) %>%
  unlist()%>%
  table

Other_Entries <- Other_Entries %>%
  mutate(DNAquality = ifelse(DNAquality=="---",NA,DNAquality),
         DNAquality = ifelse(DNAquality=="n/a",NA,DNAquality))%>%
  mutate(DNAquality = str_trim(DNAquality))%>%
  mutate(DNAquality = recode(DNAquality,
                             #yes band and yes smear 
                             'HMW band and smear' = 'HMW bands and smearing',
                             'HMW band and smearing' = 'HMW bands and smearing',
                             'HMW Bands and smear' = 'HMW bands and smearing',
                             
                             #yes band and no smear
                             
                             #no band and yes smear 
                             'no  HMW bands and smearing' = 'no HMW bands and smearing',
                             'no HMW band and smearing' = 'no HMW bands and smearing', 
                             'No HMW bands and smearing' = 'no HMW bands and smearing',

                             #no band and no smear
                             'No HMW bands and no smearing' = 'no HMW bands and no smearing',
                             'No HMW bands and no smearing' = 'no HMW bands and no smearing',
                             'no HMW bands and no smearing (empty)' = 'no HMW bands and no smearing',
                             'no HMW bands and no smearing (empty) (empty)' = 'no HMW bands and no smearing',
                             'No HMW band' = 'no HMW bands and no smearing',
                             'No HMW Bands' = 'no HMW bands and no smearing'))



##### PAUSING HERE

#SampleRoom 
Other_Entries <- Other_Entries %>%
  mutate(SampleRoom = ifelse(SampleRoom=="---",NA,SampleRoom),
         SampleRoom = ifelse(SampleRoom=="Ichthy","Ichthyology",SampleRoom),
         SampleRoom = ifelse(SampleRoom=="ichthyology","Ichthyology",SampleRoom),
         SampleRoom = ifelse(SampleRoom=="Icthology","Ichthyology",SampleRoom),
         SampleRoom = ifelse(SampleRoom=="Icthyology","Ichthyology",SampleRoom))

NoExtractionPlate <- Other_Entries %>%
  filter(is.na(ExtractionPlate))

NoExtractionPlate %>%
  filter(SampleRoom == "Ichthyology") %>%
  count(SampleLocation, SampleSublocation, DonatingLab, Collector, CommonName, Family, Genus, Species, name = "n") 

NoExtractionPlate %>%
  filter(is.na(SampleRoom)) %>%
  count(SampleLocation, SampleSublocation, DonatingLab, Collector, CommonName, Family, Genus, Species, name = "n") 


Other_Entries %>%
  filter(ExtractionPlate == "tube")




Salmon_Entries %>%
  select(ExtractionWell) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(ExtractionWell = ifelse(ExtractionWell=="---",NA,ExtractionWell))
```

what to do with these samples? confirm their physical location and then keep in database? 
what about samples with missing sample location info? 

check location of ABLG_37256 - only one sample; no plate info but sublocation box is there.... 
Kim - find location of gadid DNA and update

ask wes about species/collections with lots of missing samples:
large mouth bass and channel catfish from Cory Dick
pacific herring from Sharon Wildes
arctic cod from Kris Cieciel, Ed Farley, NOAA
numerous pollock and pcod entries... maybe this info is with metadata from Sara?
crab from Louis Copeman 
flathead sole/bering flounder/yellowfin sole from Ingrid Spies


#### Lab fields

```{r Survey, eval=T, echo=F}


Salmon_Entries %>%
  select(DNAroom) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(DNAroom = ifelse(DNAroom=="---",NA,DNAroom))

Salmon_Entries %>%
  select(DNAlocation) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(DNAlocation = ifelse(DNAlocation=="---",NA,DNAlocation))%>%
  mutate(DNAlocation = recode(DNAlocation,
                              'F212_1' = 'F212-1',
                              'F213_7' = 'F213-7'))


Salmon_Entries %>%
  select(DNAsublocation) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(DNAsublocation = ifelse(DNAsublocation=="---",NA,DNAsublocation))

#WGS1
Salmon_Entries %>%
  select(WGSDateYYYYMMDD) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(WGSBatchID) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(WGSLaneID) %>%
  unlist()%>%
  table

#WGS2 -None have been used multiple times, all blanks
Salmon_Entries %>%
  select(WGSDate2YYYYMMDD) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(WGSBatchID2) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(WGSLaneID2) %>%
  unlist()%>%
  table

#GTseq - All left empty... just not updated. 
Salmon_Entries %>%
  select(GTSeqDateYYYYMMDD) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(GTSeqPanel) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(GTSeqDate2YYYYMMDD) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(GTSeqPanel2) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(GTSeqDate3YYYYMMDD) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(GTSeqPanel3) %>%
  unlist()%>%
  table

#Genetic ID missing

#Comments are fine?

```






####### resume working here ######## 



### Contact, Donating Lab, Collector

In the field descriptions 'Contact' is listed as a required field with 
instructions about field validations; however there are values listed in
the sheet. Additionally, contact is not listed in the 'ABLG-DB-working-copy'
sheet for upload. So perhaps that field has been replaced with 
'Collector' which has 69 unique values. 

```{r StndrzCollector, eval=T, echo=T}
Salmon_Entries %>%
  select(Collector) %>%
  table
  
Salmon_Entries <- Salmon_Entries %>%
  mutate(Collector = recode(Collector,
                            'LPW-Waters' = 'Charlie Waters',
                            'Vulstek' = 'Scott Vulstek',
                            'Mike-Courtney' = 'Michael Courtney'))%>%
  mutate(Collector = ifelse(Collector=="---",NA,Collector))

NoCollector <- Salmon_Entries %>%
  filter(is.na(Collector)) 

```

There are 10,941 records without a collector recorded. Maybe that is fine provided
we have a contact/donating lab for the sample. 

Now take a stab at the Donating Lab

```{r StndrzCollector, eval=T, echo=T}
Salmon_Entries %>%
  select(DonatingLab) %>%
  table
  
Salmon_Entries <- Salmon_Entries %>%
  mutate(DonatingLab = recode(DonatingLab,
                            'LPW/Waters' = 'AFSC-LPW',
                            'Little Port Walter' = 'AFSC-LPW',
                            'Vulstek' = 'AFSC-Scott Vulstek',
                            'Vulstek?' = 'AFSC-Scott Vulstek',
                            'Seitz' = 'UAF-Andy Seitz',
                            'Manhard Mol Bio 2018' = 'UAF-Anthony Gharrett'))%>%
  mutate(DonatingLab = ifelse(DonatingLab=="---",NA,DonatingLab))

Salmon_Entries %>%
  select(DonatingLab) %>%
  table

sum(is.na(Salmon_Entries$DonatingLab))

```

```{r UnknwnRcd, eval=T, echo=T}
BadRcrds <- Salmon_Entries %>%
  filter(is.na(DonatingLab)&is.na(Collector))
```

What are the samples from 

Sckye_2014_Tog_XX - Little Togiak Beach
PL_XX - Sckye_LkWA2001
11/7/00_CedarSockeye_XX - Sckye_Cedar_1999_2000
Cedar_R_1999_ - 
Sckye2001_Baker_

BC22_25300_5006_des/env Chinook bycatch?

MCGL_18_ - Coregonus clupeaformis? whitefish_01 plate; Jared Homola Molecular Conservation Genetics Lab

DipacMale/FemaleCoho_2011; 

```{r}
Salmon_Entries <- Salmon_Entries %>%
  mutate(DonatingLab = ifelse(is.na(DonatingLab) & CommonName == 'lake whitefish','MCGL-Jared Homola',DonatingLab))
```



#### Collection Year, Month, Day

```{r Year, eval=T, echo=F}
Salmon_Entries %>%
  select(CollectionYear) %>%
  unlist()%>%
  table
  
NoYear <- Salmon_Entries %>%
  filter(is.na(CollectionYear))
```

```{r Month, eval=T, echo=F}
Salmon_Entries %>%
  select(CollectionMonth) %>%
  unlist()%>%
  table
 
Salmon_Entries %>%
  filter(CollectionMonth == 'NA')

Salmon_Entries <- Salmon_Entries %>%
  mutate(CollectionMonth = ifelse(CollectionMonth=="NA",NA,CollectionMonth))

NoMonth <- Salmon_Entries %>%
  filter(is.na(CollectionMonth))
```

```{r Year, eval=T, echo=F}
Salmon_Entries %>%
  select(CollectionDay) %>%
  unlist()%>%
  table
  
ShitSamp <- Salmon_Entries %>%
  filter(CollectionDay=='103') %>%
  select(ABLG,AlternateID,CommonName,Genus,Species,CollectionYear,CollectionMonth,CollectionDay,Location1,Location2,FreshwaterWatershed)

Salmon_Entries %>%
  filter(ABLG %in% c(41646,41647,41648,41649,41650)) %>%
  select(ABLG,AlternateID,CommonName,Genus,Species,CollectionYear,CollectionMonth,CollectionDay,Location1,Location2,FreshwaterWatershed)

#Turns out that Little Whitefish should be 8/11
Salmon_Entries %>%
  filter(grepl(pattern='SLWHI11',AlternateID))%>%
  select(ABLG,AlternateID,CommonName,Genus,Species,CollectionYear,CollectionMonth,CollectionDay,Location1,Location2,FreshwaterWatershed)

Salmon_Entries <- Salmon_Entries %>%
  mutate(CollectionDay = ifelse(grepl(pattern='SLWHI11',AlternateID),'11',CollectionDay))
  
Salmon_Entries <- Salmon_Entries %>%
  mutate(CollectionDay = recode(CollectionDay,
                                '1' = '01',
                                '2' = '02',
                                '3' = '03',
                                '4' = '04',
                                '5' = '05',
                                '6' = '06',
                                '7' = '07',
                                '8' = '08',
                                '9' = '09'))

Salmon_Entries %>%
  select(CollectionDay) %>%
  unlist()%>%
  table
```


### Location

```{r MR, eval=T, echo=F}
Salmon_Entries %>%
  select(MarineRegion) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(MarineRegion = ifelse(MarineRegion=="---",NA,MarineRegion))

```

```{r FWWS, eval=T, echo=F}
Salmon_Entries %>%
  select(FreshwaterWatershed) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(FreshwaterWatershed = ifelse(FreshwaterWatershed=="---",NA,FreshwaterWatershed))

```

```{r Loc1, eval=T, echo=F}
Salmon_Entries %>%
  select(Location1) %>%
  unlist()%>%
  table

Chickamin <- Salmon_Entries %>%
  filter(grepl(pattern='CHICKAMIN',Location1))%>%
  select(ABLG,AlternateID,CommonName,Genus,Species,CollectionYear,CollectionMonth,CollectionDay,Location1,Location2,FreshwaterWatershed)

Salmon_Entries <- Salmon_Entries %>%
  mutate(Location1 = gsub(pattern='"Replace with',replacement="",Location1))%>%
  mutate(Location1 = gsub(pattern='CHICKAMIN',replacement="Chickamin",Location1))%>%
  mutate(Location1 = gsub(pattern='ANDREW',replacement="Andrew",Location1))%>%
  mutate(Location1 = ifelse(Location1=="---",NA,Location1))%>%
  mutate(Location1 = str_trim(Location1))

```

```{r Loc2, eval=T, echo=F}
Salmon_Entries %>%
  select(Location2) %>%
  unlist()%>%
  table

#Probably below the wier samples
AukeFalls <- Salmon_Entries %>%
  filter(grepl(pattern='Auke Cr. Falls',Location2))%>%
  select(ABLG,AlternateID,CommonName,Genus,Species,CollectionYear,CollectionMonth,CollectionDay,Location1,Location2,FreshwaterWatershed)

Salmon_Entries <- Salmon_Entries %>%
  mutate(Location2 = ifelse(Location2=="---",NA,Location2))%>%
  mutate(Location2 = str_trim(Location2))

```


```{r Lat, eval=T, echo=F}
Salmon_Entries %>%
  select(StartLatitudeDD) %>%
  unlist()%>%
  table

#0 isn't valid
Salmon_Entries <- Salmon_Entries %>%
  mutate(StartLatitudeDD = ifelse(StartLatitudeDD==0,NA,StartLatitudeDD))

Salmon_Entries %>%
  select(EndLatitudeDD) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(EndLatitudeDD = ifelse(EndLatitudeDD=="---",NA,EndLatitudeDD))
```

```{r Long, eval=T, echo=F}
Salmon_Entries %>%
  select(StartLongitudeDD) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(EndLongitudeDD) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(EndLongitudeDD = ifelse(EndLongitudeDD=="---",NA,EndLongitudeDD))
```

```{r Long, eval=T, echo=F}
Salmon_Entries %>%
  select(EstimatedLatLongYesNo) %>%
  unlist()%>%
  table
```

The estimated or not estimated lat/long for each collection should be 
entered with the metadata, looks like I might have to go back and 
fill these in for salmon collections. 

### Survey Specific Fields maybe not super relevant to salmon

```{r Survey, eval=T, echo=F}
Salmon_Entries %>%
  select(CruiseID) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(HaulID) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(StationID) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(Equipment_Gear) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(DepthM) %>%
  unlist()%>%
  table
#is 0 valid for depth in meters?

Salmon_Entries %>%
  select(StartDepthM) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(EndDepthM) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(TempC) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(Habitat) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(CruiseID = ifelse(CruiseID=="---",NA,CruiseID))%>%
  mutate(HaulID = ifelse(HaulID=="---",NA,HaulID))%>%
   mutate(StationID = ifelse(StationID=="---",NA,StationID))%>%
  mutate(Habitat = ifelse(Habitat=="---",NA,Habitat))%>%
   mutate(Equipment_Gear = ifelse(Equipment_Gear=="---",NA,Equipment_Gear))%>%
  mutate(DepthM = ifelse(DepthM==0,NA,DepthM))%>%
  mutate(TempC = ifelse(TempC==0,NA,TempC))

```

#### Morphometrics

```{r Survey, eval=T, echo=F}
Salmon_Entries %>%
  select(LengthMM) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(LengthMM = ifelse(LengthMM==0,NA,LengthMM))

Salmon_Entries %>%
  select(LengthType) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(FishDepthMM) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  filter(!is.na(FishDepthMM)) %>%
select(ABLG,AlternateID,CommonName,Genus,Species,CollectionYear,CollectionMonth,CollectionDay,Location1,Location2,FreshwaterWatershed)


Salmon_Entries %>%
  filter(!is.na(FishDepthMM)) %>%
select(ABLG,AlternateID,CommonName,CollectionYear,CollectionMonth,Location1,FishDepthMM)


Salmon_Entries %>%
  select(WeightG) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(WeightG = ifelse(WeightG==0,NA,WeightG))

#That 17,940g fish is 39.5lbs!


Salmon_Entries %>%
  select(Lifestage) %>%
  unlist()%>%
  table

#These are all ages, what the hell is 1.10000000001, 2.20000002, 2.29999998. Need to go back to the actual data for those samples
# should this be more of a categorical lifestage not age? smolt, juvenile, adult etc? We have age in years
Salmon_Entries <- Salmon_Entries %>%
  mutate(Lifestage = ifelse(Lifestage=="---",NA,Lifestage))

StrngLS <- Salmon_Entries %>%
  filter(Lifestage %in% c("1.1000000000000001","2.2000000000000002","2.2999999999999998"))%>%
  select(ABLG,AlternateID,CommonName,CollectionYear,CollectionMonth,Location1,Lifestage,AgeYRS)

Salmon_Entries %>%
  select(AgeYRS) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(AgeYRS = ifelse(AgeYRS=="---",NA,AgeYRS))

Salmon_Entries %>%
  select(Sex) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(Sex = ifelse(Sex=="---",NA,Sex))%>%
  
Salmon_Entries %>%
  select(CaptureTime) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(ReleaseTime) %>%
  unlist()%>%
  table

Salmon_Entries %>%
  select(IdentifiedBy) %>%
  unlist()%>%
  table

Salmon_Entries <- Salmon_Entries %>%
  mutate(IdentifiedBy = ifelse(IdentifiedBy=="---",NA,IdentifiedBy))

```


